---
title: alm tutorial
package_version: 0.4.0
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/alm/",
	fig.width = 6,
	fig.height = 4,
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE
)
```

## What are article level metrics?

Glad you asked. The canonical URL for this is perhaps [altmetrics.org](http://altmetrics.org/manifesto/). Basically it is a metric that measures something about an article. This is in stark contrast to journal level metrics, like the Journal Impact Factor.

## Are there other altmetrics data providers?

Yes indeedy.

+ [ImpactStory](http://impactstory.it/)
+ [Altmetric.com](http://altmetric.com/)
+ [PlumAnalytics](http://www.plumanalytics.com/)

## Authentication

You are required to use an API key to access any Lagoota ALM API.

Get your API key for each service at:

* [PLOS](http://alm.plos.org/)
* [Crossref](http://det.labs.crossref.org/)
* [eLife](http://lagotto.svr.elifesciences.org/)
* [PKP](http://pkp-alm.lib.sfu.ca/)
* [Pensoft](http://alm.pensoft.net:81/)

Put your API key in your `.Rprofile` file using exactly this:
options(PlosApiKey = "YOUalmAPIKEY"),
and the functions within this package will be able to use your API key without you having to enter it every time you run a search.

Or just pss in your key using the `key` parameter in function calls. Or keep your key as a system variable, named `ALM_KEY`.

Of course you need a different key for each Lagotta instance (e.g., what works for PLOS does not work for Crossref). So just pass in your key with the `key` parameter. You can still store your keys in options, but just pass them on in the function call.

## Help

If you are having trouble with this R package, contact [the maintainer, Scott](mailto:myrmecocystus@gmail.com). If you are having trouble with the API itself, there is a newish discussion forum for the Lagotto service at [http://discuss.lagotto.io/](http://discuss.lagotto.io/).

<section id="installation">

## Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("alm")
```

Or development version from GitHub

```{r install, eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/alm")
```

Load `alm`

```{r}
library(alm)
```

<section id="usage">

## Usage

## Get data

```{r alm1}
alm_ids(doi="10.1371/journal.pone.0029797")
```

## Details for a single DOI

```{r alm2}
out <- alm_ids(doi='10.1371/journal.pone.0029797', info='detail')
```

info

```{r}
out$data$info
```

sign posts

```{r}
out$data$signposts
```

totals

```{r}
out$data$totals
```

history

```{r}
out$data$sum_metrics
```

## Search using various identifiers, including pubmed id, pmc id, and mendeley id

A single PubMed ID (pmid)

```{r}
alm_ids(pmid=22590526)
```

A single PubMed Central ID (pmcid)

```{r}
alm_ids(pmcid=212692)
```

## Search on many identifiers

```{r}
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117',
	'10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
out <- alm_ids(doi=dois)
lapply(out$data, head)
```

## Get altmetrics by year

You can also get metrics by day (`sum_metrics='day'`) or month (`sum_metrics='month'`)

```{r}
alm_ids(doi='10.1371/journal.pone.0036240', sum_metrics='year')
```

## Output an-easy-to-combine-with-other-results data.frame

```{r}
alm_ids(doi='10.1371/journal.pone.0035869', total_details=TRUE)$data[,3:7]
```

## Get detailed data for altmetrics using `almevents`

```{r}
out <- alm_events(doi="10.1371/journal.pone.0029797")
names(out)
```

Get the results for PubMed Central

```{r}
# remove those with no data
out <- out[!out %in% c("sorry, no events content yet","parser not written yet")]
out[["pmc"]]
```

Get the results for twitter

```{r}
out$twitter$events[1:3,]
```

## Retrieve and plot PLOS article-level metrics signposts.

```{r fig.keep='high'}
dat <- alm_signposts(doi="10.1371/journal.pone.0029797")
plot_signposts(input=dat)
```

Or plot many identifiers gives a line chart

```{r fig.width=7}
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117','10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
dat <- alm_signposts(doi=dois)
plot_signposts(input=dat) + theme_grey(base_size = 12)
```

Or make an interactive chart by doing `plot_signposts(input=dat, type="multiBarChart")`. Try it out! It should open in your browser and you can interact with it.

## Density and histogram plots from PLOS Article Level Metrics data

Note: Do you the key below in the `searchplos` call in this example, but if you plan to use rplos more, get your own API key [here](http://api.plos.org/).

```{r}
library('rplos')
library('plyr')
dois <- searchplos(q='science', fl="id", fq=list('cross_published_journal_key:PLoSONE', 'doc_type:full', 'publication_date:[2010-01-01T00:00:00Z TO 2010-12-31T23:59:59Z]'), limit=75)
```

Remove non-full article DOIs

```{r}
dois <- dois$data$id
dois <- dois[!grepl("annotation", dois)]
```

Collect altmetrics data and combine to a `data.frame` with `ldply`

```{r}
alm <- alm_ids(doi=dois, total_details=TRUE)
alm <- ldply(alm$data)
```

The default plot

```{r fig.width=5, fig.height=4}
plot_density(alm)
```

You can change the color of the density plot

```{r fig.width=5, fig.height=4}
plot_density(alm, color="#EFA5A5")
```

Pass in a title or description subtending the title

```{r fig.width=5, fig.height=4}
plot_density(alm, title="Scopus citations from 2010")
```

Plot a particular source

```{r fig.width=5, fig.height=4}
plot_density(alm, source="crossref_total")
```

Plot many sources in different panels in the same plot, and pass in colors just for fun

```{r fig.width=7, fig.height=6}
plot_density(input=alm, source=c("counter_html","crossref_total","pubmed_total","scopus_total"), color=c("#83DFB4","#EFA5A5","#CFD470","#B2C9E4")) + theme_grey(base_size = 12)
```

## CrossRef article level metrics

Remember to get your api key from CrossRef, pass it in in the key parameter. Notice that we are passing the base url for the Crossref API, whereas the default is for the PLOS url [http://alm.plos.org/api/v3/articles](http://alm.plos.org/api/v3/articles).

```{r}
url <- "http://alm.labs.crossref.org/api/v5/articles"
dois <- c("10.1371/journal.pone.0086859", "10.1038/nature12990", "10.5860/choice.51-3037")
alm_ids(doi=dois, url = url, key=getOption("crossrefalmkey"))
```

<section id="citing">

## Citing

To cite `alm` in publications use:

<br>

> Scott Chamberlain, Carl Boettiger, Karthik Ram and Fenner Martin (2014). alm: R Client for the Lagotto Altmetrics API Platform.. R package version 0.3.1. https://github.com/ropensci/alm

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for alm](https://github.com/ropensci/alm/issues?state=open)

[Back to top](#top)
