---
title: binomen tutorial
package_version: 0.1.0
---

```{r, eval=TRUE, echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`binomen` provides various taxonomic classes for defining a single taxon, multiple taxa, and a taxonomic data.frame

It is sort of a companion to [taxize](https://github.com/ropensci/taxize), where you can get taxonomic data on taxonomic names from the web.

The classes (S3):

* `taxon`
* `taxonref`
* `taxonrefs`
* `binomial`
* `grouping` (i.e., classification)

The verbs:

* `gethier()` - get hierarchy from a `taxon` class
* `scatter()` - make each row in taxonomic data.frame (`taxondf`) a separate `taxon` object within a single `taxa` object
* `assemble()` - make a `taxa` object into a `taxondf` data.frame
* `pick()` - pick out one or more taxonomic groups
* `pop()` - pop out (drop) one or more taxonomic groups
* `span()` - pick a range between two taxonomic groups (inclusive)
* `strain()` - filter by taxonomic groups, like dplyr's filter
* `name()` - get the taxon name for each `taxonref` object
* `uri()` - get the reference uri for each `taxonref` object
* `rank()` - get the taxonomic rank for each `taxonref` object
* `id()` - get the reference uri for each `taxonref` object


### Installation

Install and load `binomen` into the R session. Stable version from CRAN

```{r eval=FALSE}
install.packages("binomen")
```

Or development version from Github:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/binomen")
```

```{r load}
library('binomen')
```

### Usage

#### Make a taxon

Make a taxon object

```{r}
(obj <- make_taxon(genus="Poa", epithet="annua", authority="L.",
  family='Poaceae', clazz='Poales', kingdom='Plantae', variety='annua'))
```

Index to various parts of the object

The binomial

```{r}
obj$binomial
```

The authority

```{r}
obj$binomial$authority
```

The classification

```{r}
obj$grouping
```

The family

```{r}
obj$grouping$family
```

#### Subset taxon objects

Get one or more ranks via `pick()`

```{r}
obj %>% pick(family)
obj %>% pick(family, genus)
```

Drop one or more ranks via `pop()`

```{r}
obj %>% pop(family)
obj %>% pop(family, genus)
```

Get a range of ranks via `span()`

```{r}
obj %>% span(kingdom, family)
```

Extract classification as a `data.frame`

```{r}
gethier(obj)
```

#### Taxonomic data.frame's

Make one

```{r}
df <- data.frame(order = c('Asterales','Asterales','Fagales','Poales','Poales','Poales'),
  family = c('Asteraceae','Asteraceae','Fagaceae','Poaceae','Poaceae','Poaceae'),
  genus = c('Helianthus','Helianthus','Quercus','Poa','Festuca','Holodiscus'),
  stringsAsFactors = FALSE)
(df2 <- taxon_df(df))
```

Parse - get rank order via `pick()`

```{r}
df2 %>% pick(order)
```

get ranks order, family, and genus via `pick()`

```{r}
df2 %>% pick(order, family, genus)
```

get range of names via `span()`, from rank `X` to rank `Y`

```{r}
df2 %>% span(family, genus)
```

Separate each row into a `taxon` class (many `taxon` objects are a `taxa` class)

```{r output.lines=1:20}
scatter(df2)
```

And you can re-assemble a data.frame from the output of `scatter()` with `assemble()`

```{r}
out <- scatter(df2)
assemble(out)
```

### Citing

To cite `binomen` in publications use:

> Scott Chamberlain (2016). binomen: 'Taxonomic' Specification and Parsing Methods. https://github.com/ropensci/binomen

### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rnoaa](https://github.com/ropensci/binomen/issues?state=open)

[Back to top](#top)
