---
title: ckanr tutorial
package_version: 0.1.0
---

```{r, eval=TRUE, echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

ckanr is a general purpose R client for CKAN, the open source data portal platform
from the Open Knowledge Foundation.

It is meant to be as general as possible, allowing you to work with any CKAN instance.

### Installation

Install and load `ckanr` into the R session. Stable version from CRAN

```{r eval=FALSE}
install.packages("ckanr")
```

Or development version from Github:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/ckanr")
```

```{r load}
library('ckanr')
```

### Usage

Note: the default base CKAN URL is set to [http://data.techno-science.ca/](http://data.techno-science.ca/). You can change this using `ckanr_setup()`, or change the URL using the `url`
parameter in each function call.

To set one or both, run:

```{r eval=FALSE}
# restores default CKAN url to http://data.techno-science.ca/
ckanr_setup()
# Just set url
ckanr_setup(url = "http://data.techno-science.ca/")
# set url and key
ckanr_setup(url = "http://data.techno-science.ca/", key = "my-ckan-api-key")
```

#### Changes

```{r}
changes(limit = 2, as = "table")[, 1:4]
```

#### List datasets

```{r output.lines=1:10}
package_list(as = "table")
```

#### List tags

```{r}
tag_list('aviation', as = 'table')
```

#### Show tags

Subset for readme brevity

```{r}
tag_show('Aviation')
```

#### List groups

```{r}
group_list(as = 'table')[, 1:3]
```

#### Show groups

Subset for readme brevity

```{r}
group_show('communications', as = 'table')$users
```

#### Show a package

```{r}
package_show('34d60b13-1fd5-430e-b0ec-c8bc7f4841cf', as = 'table')$resources[, 1:10]
```

#### Search for packages

```{r}
out <- package_search(q = '*:*', rows = 2, as = "table")$results
out[, !names(out) %in% 'resources'][, 1:10]
```

#### Search for resources

```{r}
resource_search(q = 'name:data', limit = 2, as = 'table')
```

#### Example of using a different CKAN API

**The UK Natural History Museum**

Website: [http://data.nhm.ac.uk/](http://data.nhm.ac.uk/)

List datasets

```{r output.lines=1:10}
ckanr_setup(url = "http://data.nhm.ac.uk")
package_list(as = "table")
```

Tags

_list_

```{r}
head(tag_list(as = "table"))
```

_show_

```{r}
tag_show('arthropods', as = 'table')
```

Packages

_search_

```{r}
out <- package_search(q = '*:*', rows = 2, as = 'table')
out$results[, 1:10]
```

_show_

```{r output.lines=1:10}
package_show(id = "56e711e6-c847-4f99-915a-6894bb5c5dea", as = "table")
```

### Citing

To cite `ckanr` in publications use:

<br>

> Scott Chamberlain, Imanuel Costigan, Wush Wu, Florian Mayer (2016). ckanr: Client for the Comprehensive Knowledge Archive Network ('CKAN') 'API'. R package version 0.1.0. https://github.com/ropensci/ckanr

### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rnoaa](https://github.com/ropensci/ckanr/issues?state=open)

[Back to top](#top)
