---
title: etseed tutorial
package_version: 0.1.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`etcd` is a key-value DB written in `Go`. It has an HTTP API, which this R package wraps.

[etcd API docs](https://github.com/coreos/etcd/blob/master/Documentation/v2/api.md)

**Installing etcd**

See the [etcd Github repo](https://github.com/coreos/etcd#etcd) for help on installing `etcd`.

There are various ways to install it, and they depend on your operating sytsem.

You can install via `homebrew`, install from source, and via Docker.

**Start etcd**

at the command line

```sh
etcd
```

> how to start etcd may differ depending on your setup


### Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("etseed")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropensci/etseed")
```

```{r}
library("etseed")
```


### Usage

**Make a client**

First task when using this package is to initialize a client
with the `etcd()` function. it's a wrapper around an R6 class.

```{r}
(client <- etcd())
```

Default settings in `etcd()` connect you to `localhost`, and port `2379`,
using etcd API version 2, with an `http` scheme.

**Get version**

```{r}
client$version()
```

**Create a directory**

```{r echo=FALSE, results='hide'}
out <- tryCatch(client$key("/neighbor"), error=function(e) e)
if (!inherits(out, "http_error")) client$delete("/neighbor", dir=TRUE)
Sys.sleep(3)
```

```{r}
client$create("/neighbor", dir = TRUE)
```

**Create a key**

```{r echo=FALSE, results='hide'}
out <- tryCatch(client$key("/mykey"), error=function(e) e)
if (!inherits(out, "simpleError")) client$delete("/mykey")
out <- tryCatch(client$key("/stuff"), error=function(e) e)
if (!inherits(out, "simpleError")) client$delete("/stuff")
```

```{r}
client$create(key = "/mykey", value = "this is awesome")
```

```{r echo=FALSE, results='hide'}
Sys.sleep(3)
```

Use `ttl` parameter to make it dissappear after `x` seconds

```{r}
client$create(key = "/stuff", value = "tables", ttl = 5)
```

And the key will be gone after 5 seconds, see:

```{r eval=FALSE}
client$key("/stuff")
#> Error in etcd_GET(sprintf("%s%s/%s/", etcdbase(), "keys", key), ...) :
#>   client error: (404) Not Found
```

**Update a key**

```{r echo=FALSE, results='hide'}
out <- tryCatch(client$key("/foo"), error=function(e) e)
if (!inherits(out, "simpleError")) client$delete("/foo")
```

Create a key

```{r}
client$create(key = "/foo", value = "bar")
```

Then update the key

```{r output.lines = 1:10}
client$update(key = "/foo", value = "bar stool")
```

**Create in-order keys**

```{r}
client$create_inorder("/queue", "thing1")
```

```{r}
client$create_inorder("/queue", "thing2")
```

```{r}
client$create_inorder("/queue", "thing3")
```

**List keys**

```{r output.lines = 1:10}
client$keys()
```

**List a key**

```{r}
client$key("/mykey")
```



### Citing

> Scott Chamberlain (2016). etseed: Client for 'etcd', a 'Key-value' Database. R package version 0.1.0. https://cran.rstudio.com/package=etseed



### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for etseed](https://github.com/ropensci/etseed/issues?state=open)


[Back to top](#top)
