---
title: geojsonlint tutorial
package_version: 0.1.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/geojsonlint/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

Tools for linting 'GeoJSON'.

Package API:

* `geojson_lint()` - [GeoJSON lint web service](http://geojsonlint.com/)
* `geojson_hint()` - [GeoJSON hint JS library](https://www.npmjs.com/package/geojsonhint)
* `geojson_validate()` - [is-my-json-valid JS library](https://www.npmjs.com/package/is-my-json-valid) using a GeoJSON
schema from [https://github.com/SchemaStore/schemastore](https://github.com/SchemaStore/schemastore)

<section id="installation">

## Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("geojsonlint")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropenscilabs/geojsonlint")
```

```{r}
library("geojsonlint")
```

<section id="usage">

## Usage

### Good GeoJSON

geojsonlint.com web service

```{r eval = FALSE}
geojson_lint(x = '{"type": "Point", "coordinates": [-100, 80]}')
#> [1] TRUE
```

geojsonhint JS library

```{r}
geojson_hint(x = '{"type": "Point", "coordinates": [-100, 80]}')
```

is-my-json-valid JS library

```{r}
geojson_validate(x = '{"type": "Point", "coordinates": [-100, 80]}')
```

### Bad GeoJSON

geojsonlint.com web service

```{r eval = FALSE}
geojson_lint('{"type": "FooBar"}')
#> [1] FALSE
```

geojsonhint JS library

```{r}
geojson_hint('{"type": "FooBar"}')
```

is-my-json-valid JS library

```{r}
geojson_validate('{ "type": "FeatureCollection" }')
```

### Bad GeoJSON - with reason for failure

geojsonlint.com web service

```{r eval = FALSE}
geojson_lint('{"type": "FooBar"}', verbose = TRUE)
#> [1] FALSE
#> attr(,"errors")
#>                                 message status
#> 1 "FooBar" is not a valid GeoJSON type.  error
```

geojsonhint JS library

```{r}
geojson_hint('{"type": "FooBar"}', verbose = TRUE)
```

is-my-json-valid JS library

```{r}
geojson_validate('{ "type": "FeatureCollection" }', verbose = TRUE)
```

### Bad GeoJSON - stop on validation failure

geojsonlint.com web service

```{r eval=FALSE}
geojson_lint('{"type": "FooBar"}', error = TRUE)
#> Error: invalid GeoJSON
#>    - "FooBar" is not a valid GeoJSON type.
```

geojsonhint JS library

```{r eval=FALSE}
geojson_hint('{"type": "FooBar"}', error = TRUE)
#> Error: Line 1
#>    - The type FooBar is unknown
```

is-my-json-valid JS library

```{r eval=FALSE}
geojson_validate('{ "type": "FeatureCollection" }', error = TRUE)
#> Error: 1 error validating json:
#> 	- data: no (or more than one) schemas match
```


<section id="citing">

## Citing

> Scott Chamberlain and Andy Teucher (2016). geojsonlint: Tools for Validating
  'GeoJSON'. R package version 0.1.0 https://cran.rstudio.com/package=geojsonlint


<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for geojsonlint](https://github.com/ropenscilabs/geojsonlint/issues?state=open)


[Back to top](#top)
