---
title: getlandsat tutorial
package_version: 0.1.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/getlandsat/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`getlandsat` provides access to Landsat <https://landsat.usgs.gov> 8 metadata and images hosted on AWS S3 at <https://aws.amazon.com/public-data-sets/landsat>. The package only fetches data. It does not attempt to aid users in downstream usage, but some additional functionality may be added.

Potential users are probably anyone from scientists asking questions about biodiversity or land use change, to software developers creating tools for users to vizualize their data.

<section id="installation">

## Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("getlandsat")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropenscilabs/getlandsat")
```

```{r}
library("getlandsat")
```

<section id="usage">

## Usage


```{r echo=FALSE}
lsat_cache_delete_all()
```


### List scenes

```{r}
(res <- lsat_scenes(n_max = 10))
```

### List scene files

```{r}
lsat_scene_files(x = res$download_url[1])
```

### Get an image

Returns path to the image

```{r}
lsat_image(x = "LC80101172015002LGN00_B5.TIF")
```

Another one

```{r}
lsat_image("LC80010032014272LGN00_B10.TIF")
```

#### Caching

When requesting an image, we first check if you already have that image. If you do,
we return the path to the file. If not, we get the image, and return the file path.

```{r message=TRUE}
lsat_image(x = "LC80101172015002LGN00_B5.TIF")
```

Note the message given.

See `?lsat_cache` for cache management functions.

### Visualize

```{r}
library("raster")
x <- lsat_cache_details()
img <- raster(x[[1]]$file)
plot(img)
```


<section id="citing">

## Citing

> Scott Chamberlain (2016). getlandsat: Get Landsat 8 Data
  from Amazon Pulic Data Sets. R package version 0.1.0.
  https://cran.rstudio.com/package=getlandsat


<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for getlandsat](https://github.com/ropenscilabs/getlandsat/issues?state=open)


[Back to top](#top)
