---
title: gutenbergr tutorial
package_version: 0.1.3
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

The gutenbergr package helps you download and process public domain works from the [Project Gutenberg](https://www.gutenberg.org/) collection. This includes both tools for downloading books (and stripping header/footer information), and a complete dataset of Project Gutenberg metadata that can be used to find words of interest. Includes:

* A function `gutenberg_download()` that downloads one or more works from Project Gutenberg by ID: e.g., `gutenberg_download(84)` downloads the text of Frankenstein.
* Metadata for all Project Gutenberg works as R datasets, so that they can be searched and filtered:
  * `gutenberg_metadata` contains information about each work, pairing Gutenberg ID with title, author, language, etc
  * `gutenberg_authors` contains information about each author, such as aliases and birth/death year
  * `gutenberg_subjects` contains pairings of works with Library of Congress subjects and topics


### Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("gutenbergr")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropenscilabs/gutenbergr")
```

```{r}
library("gutenbergr")
```


### Project Gutenberg Metadata

This package contains metadata for all Project Gutenberg works as R datasets, so that you can search and filter for particular works before downloading.

The dataset `gutenberg_metadata` contains information about each work, pairing Gutenberg ID with title, author, language, etc:

```{r}
library(gutenbergr)
gutenberg_metadata
```

For example, you could find the Gutenberg ID of Wuthering Heights by doing:

```{r}
library(dplyr)

gutenberg_metadata %>%
  filter(title == "Wuthering Heights")
```

In many analyses, you may want to filter just for English works, avoid duplicates, and include only books that have text that can be downloaded. The `gutenberg_works()` function does this pre-filtering:

```{r}
gutenberg_works()
```

It also allows you to perform filtering as an argument:

```{r}
gutenberg_works(author == "Austen, Jane")

# or with a regular expression

library(stringr)
gutenberg_works(str_detect(author, "Austen"))
```

The meta-data currently in the package was last updated on **`r format(attr(gutenberg_metadata, "date_updated"), '%d %B %Y')`**.

### Downloading books by ID

The function `gutenberg_download()` downloads one or more works from Project Gutenberg based on their ID. For example, we earlier saw that "Wuthering Heights" has ID 768 (see [the URL here](https://www.gutenberg.org/ebooks/768)), so `gutenberg_download(768)` downloads this text.

```{r}
wuthering_heights <- gutenberg_download(768)

wuthering_heights
```

Notice it is returned as a tbl_df (a type of data frame) including two variables: `gutenberg_id` (useful if multiple books are returned), and a character vector of the text, one row per line. Notice that the header and footer added by Project Gutenberg (visible [here](http://www.gutenberg.org/files/768/768.txt)) have been stripped away.

Provide a vector of IDs to download multiple books. For example, to download Jane Eyre (book [1260](https://www.gutenberg.org/ebooks/1260)) along with Wuthering Heights, do:

```{r}
books <- gutenberg_download(c(768, 1260), meta_fields = "title")

books
```

Notice that the `meta_fields` argument allows us to add one or more additional fields from the `gutenberg_metadata` to the downloaded text, such as title or author.

```{r}
books %>%
  count(title)
```

### Other meta-datasets

You may want to select books based on information other than their title or author, such as their genre or topic. `gutenberg_subjects` contains pairings of works with Library of Congress subjects and topics. "lcc" means [Library of Congress Classification](https://www.loc.gov/catdir/cpso/lcco/), while "lcsh" means [Library of Congress subject headings](http://id.loc.gov/authorities/subjects.html):

```{r}
gutenberg_subjects
```

This is useful for extracting texts from a particular topic or genre, such as detective stories, or a particular character, such as Sherlock Holmes. The `gutenberg_id` column can then be used to download these texts or to link with other metadata.

```{r}
gutenberg_subjects %>%
  filter(subject == "Detective and mystery stories")

gutenberg_subjects %>%
  filter(grepl("Holmes, Sherlock", subject))
```

`gutenberg_authors` contains information about each author, such as aliases and birth/death year:

```{r}
gutenberg_authors
```

### Analysis

What's next after retrieving a book's text? Well, having the book as a data frame is especially useful for working with the [tidytext](https://github.com/juliasilge/tidytext) package for text analysis.

```{r}
library(tidytext)

words <- books %>%
  unnest_tokens(word, text)

words

word_counts <- words %>%
  anti_join(stop_words, by = "word") %>%
  count(title, word, sort = TRUE)

word_counts
```



### Citing

> David Robinson (2017). gutenbergr: Download and Process Public Domain Works
  from Project Gutenberg. R package version 0.1.3.
  https://cran.rstudio.com/package=gutenbergr



### License and bugs

* License: [GPL-2](https://opensource.org/licenses/GPL-2.0)
* Report bugs at [our GitHub repo for gutenbergr](https://github.com/ropenscilabs/gutenbergr/issues?state=open)


[Back to top](#top)
