---
title: historydata tutorial
package_version: 0.1
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.path="../assets/tutorial-images/historydata/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width=6,
  fig.height=4,
  cache=TRUE
)
```

Datasets for Historians

<section id="installation">

## Installation

```{r eval=FALSE}
install.packages("historydata")
```

Or development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/historydata")
```

```{r}
library("historydata")
```

<section id="usage">

## Usage

List all the datasets in the package with their documentation:

```{r eval=FALSE}
help(package = historydata)
```

Which opens help in GUI in Rstudio/R.app, or list within R

```{r}
ls("package:historydata")
```

To load a dataset, e.g, the dataset on Roman Catholic dioceses in the United States, Canada, and Mexico

```{r}
data("catholic_dioceses")
```

Peek at the dataset

```{r}
head(catholic_dioceses)
```

Another dataset, of the Federal judges in the United States of America

```{r}
data("judges_people")
```

```{r}
head(judges_people)
```


<section id="citing">

## Citing

To cite `historydata` in publications use:

<br>

> Lincoln Mullen (2014). historydata: Data Sets for Historians. R package version 0.1.
  https://cran.rstudio.com/package=historydata

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for historydata](https://github.com/ropensci/historydata/issues?state=open)

[Back to top](#top)
