---
title: neotoma tutorial
package_version: 1.7.0
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "img/tutorial-images/neotoma/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

The `neotoma` package is a programmatic R interface to the [Neotoma Paleoecological Database](http://www.neotomadb.org/). The package is intended to both allow users to search for sites and to download data for use in analyical workflows of paleoecological research.

For more information on the package please refer to:

Goring, S., Dawson, A., Simpson, G. L., Ram, K., Graham, R. W., Grimm, E. C., & Williams, J. W.. (2015). neotoma: A Programmatic Interface to the Neotoma Paleoecological Database. *Open Quaternary*, 1(1), Art. 2. DOI: [10.5334/oq.ab](http://doi.org/10.5334/oq.ab)


### Installation

```{r eval=FALSE}
install.packages("neotoma")
```

Or development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/neotoma")
```

```{r}
library("neotoma")
```


### Distribution of sites with Mammoth fossils in Neotoma

Example requires the mapdata package:

```{r}
library('mapdata')
```

You may use either '%' or "*" as wildcards for search terms:


```{r}
test <- get_dataset(taxonname='Mammuthus*')
```

The API call was successful, you have returned  3273 records.

```{r}
site.locs <- get_site(test)
```

A crude way of making the oceans blue.

```{r}
plot(1, type = 'n',
     xlim=range(site.locs$long)+c(-10, 10),
     ylim=range(site.locs$lat)+c(-10, 10),
     xlab='Longitude', ylab = 'Latitude')
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = "lightblue")
map('world',
    interior=TRUE,
    fill=TRUE,
    col='gray',
    xlim=range(site.locs$long)+c(-10, 10),
    ylim=range(site.locs$lat)+c(-10, 10),
    add=TRUE)
points(site.locs$long, site.locs$lat, pch=19, cex=0.5, col='red')
```

### Proportion of publications/year for Neotoma datasets

```{r}
library('ggplot2')
library('plyr')
pubs <- get_publication()
pub.years <- ldply(pubs, "[[", "meta")
ggplot(pub.years, aes(x = year)) +
     stat_bin(aes(y = ..density..*100, position = 'dodge'), binwidth = 1) +
     theme_bw() +
     ylab('Percent of Publications') +
     xlab('Year of Publication') +
     scale_y_continuous(expand = c(0, 0.1)) +
     scale_x_continuous(
       breaks = seq(min(pub.years$year, na.rm = TRUE), 2014, by = 20))
```



### Citing

To cite `neotoma` in publications use:

<br>

> Goring, S., Dawson, A., Simpson, G. L., Ram, K., Graham, R. W.,
  Grimm, E. C., & Williams, J. W.. (2015). neotoma: A Programmatic
  Interface to the Neotoma Paleoecological Database, 1(1), Art. 2. DOI:
  http://doi.org/10.5334/oq.ab


### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for neotoma](https://github.com/ropensci/neotoma/issues?state=open)

[Back to top](#top)
