---
title: oai vignette
package_version: 0.2.2
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "img//utorial-images/oai/",
	warning = FALSE,
	message = FALSE,
	comment = "#>"
)
library("oai", quietly = TRUE)
```

A general purpose client to work with any 'OAI-PMH' service. The 'OAI-PMH' protocol is described at [http://www.openarchives.org/OAI/openarchivesprotocol.html](http://www.openarchives.org/OAI/openarchivesprotocol.html). The main functions follow the OAI-PMH verbs:

* `GetRecord`
* `Identify`
* `ListIdentifiers`
* `ListMetadataFormats`
* `ListRecords`
* `ListSets`

### Installation

You can install from CRAN

```{r eval=FALSE}
install.packages("oai")
```

Or the development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/oai")
```


### Identify

```{r}
id("http://oai.datacite.org/oai")
```

### ListIdentifiers

```{r}
list_identifiers(from = '2011-05-01T', until = '2011-09-01T')
```

### Count Identifiers

```{r}
count_identifiers()
```

### ListRecords

```{r}
list_records(from = '2011-05-01T', until = '2011-08-15T')
```

### GetRecords

```{r}
get_records(c("oai:oai.datacite.org:32255", "oai:oai.datacite.org:32325"))
```

### List MetadataFormats

```{r}
list_metadataformats(id = "oai:oai.datacite.org:32348")
```

### List Sets

```{r}
list_sets("http://oai.datacite.org/oai")
```

### Examples of other OAI providers

#### Global Biodiversity Information Facility

Identify

```{r}
id("http://api.gbif.org/v1/oai-pmh/registry")
```

Get records

```{r}
get_records(c("816f4734-6b49-41ab-8a1d-1b21e6b5486d", "95e3042f-f48d-4a04-8251-f755bebeced6"),
            url = "http://api.gbif.org/v1/oai-pmh/registry")
```

#### Biodiversity Heritage Library

Identify

```{r}
id("http://www.biodiversitylibrary.org/oai")
```

Get records

```{r}
get_records(c("oai:biodiversitylibrary.org:item/7", "oai:biodiversitylibrary.org:item/9"),
            url = "http://www.biodiversitylibrary.org/oai")
```


### Citing

> Scott Chamberlain and Michal Bojanowski (2016). oai: General Purpose 'Oai-PMH' Services Client. R package
  version 0.2.2. https://github.com/ropensci/oai




### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for oai](https://github.com/ropensci/oai/issues?state=open)


[Back to top](#top)
