---
title: originr tutorial
package_version: 0.2.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "img/tutorial-images/originr/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

Species Origin Data

Data sources:

* Encyclopedia of Life [http://eol.org](http://eol.org)
* Flora Europaea [http://rbg-web2.rbge.org.uk/FE/fe.html](http://rbg-web2.rbge.org.uk/FE/fe.html)
* Global Invasive Species Database [http://www.iucngisd.org/gisd](http://www.iucngisd.org/gisd)
* Native Species Resolver [http://bien.nceas.ucsb.edu/bien/tools/nsr/nsr-ws/](http://bien.nceas.ucsb.edu/bien/tools/nsr/nsr-ws/)
* Integrated Taxonomic Information Service [http://www.itis.gov/](http://www.itis.gov/)



### Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("originr")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropenscilabs/originr")
```

```{r}
library("originr")
```


## Usage

### EOL invasive species datasets

Datasets included:

* `gisd100` - 100 of the World's Worst Invasive Alien Species
(Global Invasive Species Database) [http://eol.org/collections/54500](http://eol.org/collections/54500)
* `gisd` - Global Invasive Species Database 2013 [http://eol.org/collections/54983](http://eol.org/collections/54983)
* `isc` - Centre for Agriculture and Biosciences International Invasive Species
Compendium (ISC) [http://eol.org/collections/55180](http://eol.org/collections/55180)
* `daisie` - Delivering Alien Invasive Species Inventories for Europe (DAISIE) Species
List [http://eol.org/collections/55179](http://eol.org/collections/55179)
* `i3n` - IABIN Invasives Information Network (I3N) Species
http://eol.org/collections/55176
* `mineps` - Marine Invaders of the NE Pacific Species [http://eol.org/collections/55331](http://eol.org/collections/55331)

An example using `mineps`

```{r}
eol(name='Ciona intestinalis', dataset='mineps')
```

### Native Species Resolver

```{r}
nsr("Pinus ponderosa", "United States")
```

### Global Invasive Species Database

```{r}
sp <- c("Carpobrotus edulis", "Rosmarinus officinalis")
gisd(sp)
```

### Flora Europaea

```{r}
flora_europaea("Lavandula stoechas")
```



### Citing

> Scott Chamberlain and Ignasi Bartomeus (2016). originr: Fetch Species
  Origin Data from the Web. R package version 0.2.0. https://cran.rstudio.com/package=originr



### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for originr](https://github.com/ropenscilabs/originr/issues?state=open)


[Back to top](#top)

