---
title: pangaear tutorial
package_version: 0.2.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "../assets/tutorial-images/pangaear/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

Tools to interact with the [Pangaea Database](https:// www.pangaea.de), including functions for searching for data, fetching datasets by dataset ID, and working with the Pangaea OAI-PMH service.

<section id="installation">

## Installation

Stable `pangaear` version from CRAN

```{r eval=FALSE}
install.packages("pangaear")
```

Or, the development version from Github

```{r eval=FALSE}
devtools::install_github("ropensci/pangaear")
```

```{r}
library("pangaear")
```

<section id="usage">

## Usage

### Search for data

This is a thin wrapper around the GUI search interface on the page [https://www.pangaea.de/](https://www.pangaea.de/). Everything you can do there, you can do here.

```{r}
pg_search(query = 'water', bbox = c(-124.2, 41.8, -116.8, 46.1), count = 3)
```

### Get data

```{r}
res <- pg_data(doi = '10.1594/PANGAEA.761032')
res[[1]]
```

Search for data then pass DOI to data function.

```{r}
res <- pg_search(query = 'water', bbox = c(-124.2, 41.8, -116.8, 46.1), count = 3)
pg_data(res$doi[1])
```

### OAI-PMH metadata

#### Identify the service

```{r}
pg_identify()
```

#### List metadata formats

```{r}
pg_list_metadata_formats()
```

#### List identifiers

```{r}
pg_list_identifiers(from = '2015-09-01', until = '2015-09-05')
```

#### List sets

```{r}
pg_list_sets()
```

#### List records

```{r}
pg_list_records(from = '2015-09-01', until = '2015-09-10')
```

#### Get a record

```{r}
pg_get_record(identifier = "oai:pangaea.de:doi:10.1594/PANGAEA.788382")
```


<section id="citing">

## Citing

To cite `pangaear` in publications use:

<br>

> Scott Chamberlain, Kara Woo, Andrew MacDonald, Naupaka Zimmerman and Gavin Simpson (2016). pangaear:
  Client for the 'Pangaea' Database. R package version 0.2.0. https://github.com/ropensci/pangaear

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for pangaear](https://github.com/ropensci/pangaear/issues?state=open)

[Back to top](#top)
