---
title: rAltmetric tutorial
package_version: 0.3
---

```{r, eval=TRUE, echo=FALSE}
opts_chunk$set(fig.path="../assets/tutorial-images/rAltmetric/")
```

This package provides a way to programmatically retrieve altmetric data from [altmetric.com](http://altmetric.com) for any publication with the appropriate identifer. The package is really simple to use and only has two major functions: One (`altmetrics()`) to download metrics and another (`altmetric_data()`) to extract the data into a `data.frame`. It also includes generic S3 methods to plot/print metrics for any altmetric object.

Questions, features requests and issues should go [here](https://github.com/ropensci/rAltmetric/issues/). General comments to [karthik.ram@gmail.com](mailto:karthik.ram@gmail.com).

<section id="installation">

## Installation


A stable version is available from CRAN. To install

```{r install, message=FALSE, warning=FALSE, eval=FALSE}
install.packages('rAltmetric')
```

### Development version

```{r installcran, message=FALSE, warning=FALSE, eval=FALSE}
# If you don't already have the devtools library, first run
install.packages('devtools')
```

Then install the package

```{r installgithub, message=FALSE, warning=FALSE, eval=FALSE}
library(devtools)
install_github('rAltmetric', 'ropensci')
```

<section id="usage">

## Usage


### Obtaining metrics

There was a recent paper by [Acuna et al](http://www.nature.com/news/2010/100616/full/465860a.html) that received a lot of attention on Twitter. What was the impact of that paper?

```{r one, message=FALSE, warning=FALSE}
library(rAltmetric)
```

```{r two, message=FALSE, warning=FALSE}
acuna <- altmetrics(doi="10.1038/489201a")
acuna
```

### Data

To obtain the metrics in tabular form for further processing, run any object of class `altmetric` through `altmetric_data()` to get data that can easily be written to disk as a spreadsheet.

```{r x, message=FALSE, warning=FALSE}
altmetric_data(acuna)
```

You can save these data into a clean spreadsheet format:

```{r write, message=FALSE, warning=FALSE, eval=FALSE}
acuna_data <- altmetric_data(acuna)
write.csv(acuna_data, file = 'acuna_altmetrics.csv')
```

### Visualization

For any altmetric object you can quickly plot the stats with a generic `plot` function. The plot overlays the [altmetric badge and the score](http://api.altmetric.com/embeds.html) on the top right corner. If you prefer a customized plot, create your own with the raw data generated from `almetric_data()`

```{r plot, message=FALSE, warning=FALSE}
plot(acuna)
```

### Gathering metrics for many DOIs

For a real world use-case, one might want to get metrics on multiple publications. If so, just read them from a spreadsheet and `llply` through them like the example below.

Install `rplos` (v0.3.6) if you don't have it. See [here](https://github.com/ropensci/rplos) for instructions.

```{r gather1, message=FALSE, warning=FALSE}
library(rplos)
dois <- searchplos(q = "*:*", fl = 'id', fq = 'doc_type:full', start = 0, limit = 75)$id
```

First, let's retrieve the metrics.

```{r gather2, message=FALSE, warning=FALSE}
library(plyr)
raw_metrics <- lapply(dois, function(x) altmetrics(doi = x))
```

Now let's pull the data together.

```{r gather3, message=FALSE, warning=FALSE}
metric_data <- ldply(raw_metrics, altmetric_data)
head(metric_data[,c(1:10)])
```

Finally we save this to a spreadsheet for further analysis/vizualization.

```{r gather4, message=FALSE, warning=FALSE}
write.csv(metric_data, file = "metric_data.csv")
```

### Further reading
* [Metrics: Do metrics matter?](http://www.nature.com/news/2010/100616/full/465860a.html)
* [The altmetrics manifesto](http://altmetrics.org/manifesto/)


<section id="citing">

## Citing

To cite `rAltmetric` in publications use:

<br>

> Karthik Ram (2012). rAltmetric: Retrieves altmerics data for any published paper from altmetric.com. R package version 0.5. https://github.com/ropensci/rAltmetric

<section id="license_bugs">

## License and bugs

* License: [CC0](http://creativecommons.org/choose/zero/)
* Report bugs at [our Github repo for rAltmetric](https://github.com/ropensci/rAltmetric/issues?state=open)

[Back to top](#top)
