---
title: rAltmetric tutorial
package_version: 0.7.0
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
	fig.path="img/tutorial-images/rAltmetric/",
	comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

This package provides a way to programmatically retrieve altmetric data from [altmetric.com](http://altmetric.com) for any publication with the appropriate identifer. The package is really simple to use and only has two major functions: One (`altmetrics()`) to download metrics and another (`altmetric_data()`) to extract the data into a `data.frame`. It also includes generic S3 methods to plot/print metrics for any altmetric object.

Questions, features requests and issues should go [here](https://github.com/ropensci/rAltmetric/issues/). General comments to [karthik.ram@gmail.com](mailto:karthik.ram@gmail.com).


### Installation


A stable version is available from CRAN. To install

```{r install, message=FALSE, warning=FALSE, eval=FALSE}
install.packages('rAltmetric')
```

```{r installcran, message=FALSE, warning=FALSE, eval=FALSE}
install.packages('devtools')
devtools::install_github('ropensci/rAltmetric')
```

### Usage

```{r}
library(rAltmetric)
library(magrittr)
library(purrr)

ids <- list(
  "10.1038/nature09210",
  "10.1126/science.1187820",
  "10.1016/j.tree.2011.01.009",
  "10.1086/664183"
)

alm <- function(x)  altmetrics(doi = x) %>% altmetric_data()

results <- map_df(ids, alm)
# This results in a data.frame with one row per identifier.
```

You can now see some citation data for these papers.

```{r}
library(dplyr)
results %>% select(title, doi,  starts_with("cited"))
```


### Citing

To cite `rAltmetric` in publications use:

<br>

> Karthik Ram (2017). rAltmetric: Retrieves Altmerics Data for Any
  Published Paper from 'Altmetric.com'. R package version 0.7.0.
  https://CRAN.R-project.org/package=rAltmetric


### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rAltmetric](https://github.com/ropensci/rAltmetric/issues?state=open)

[Back to top](#top)
