---
title: rbhl tutorial
package_version: 0.8.0
---

```{r, eval=TRUE, echo=FALSE}
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
})
knitr::opts_chunk$set(
  fig.path = "img/tutorial-images/rbhl/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`rbhl` is an R interface to the Biodiversity Heritage Library.


### Installation

```{r eval=FALSE}
install.packages("rbhl")
```

Or development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rbhl")
```

```{r}
library("rbhl")
```

## Usage

### Output formats

You can output various formats using the `as` parameter, setting to 'json', 'xml', 'list', or 'table'. Note that not all options are available in all functions due to varying returned data objects.

Raw output, in xml format

```{r authorsearch1, messages=FALSE, warning=FALSE, eval=FALSE}
bhl_authorsearch(name='dimmock', as='xml')
```

json format

```{r authorsearch2, messages=FALSE, warning=FALSE}
bhl_authorsearch(name='dimmock', as='json')
```

Or get a list

```{r output.lines=10}
bhl_authorsearch(name='dimmock', as='list')
```

The default option in most functions is to parse to a data.frame

```{r authorsearch4, messages=FALSE, warning=FALSE}
bhl_authorsearch(name='dimmock')
```

### Some examples of function calls

#### Get title metadata

```{r output.lines=10}
bhl_gettitlemetadata(1726, items = TRUE)
```

#### Book search

```{r booksearch, messages=FALSE, warning=FALSE}
bhl_booksearch(title='Selborne', lname='White', volume=2, edition='new', year=1825, collectionid=4, language='eng')
```

#### Search titles

```{r}
bhl_titlesearchsimple('husbandry')
```

#### Get languages

This function gets a list of languages in which materials in BHL have been written.

```{r}
bhl_getlanguages()
```

### Citing

To cite `rbhl` in publications use:

<br>

> Scott Chamberlain (2017). rbhl: Interface to the 'Biodiversity'
  'Heritage' Library. R package version 0.8.0.
  https://CRAN.R-project.org/package=rbhl

### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rbhl](https://github.com/ropensci/rbhl/issues?state=open)

[Back to top](#top)
