---
title: rbison tutorial
package_version: 0.4.8
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
	fig.path="../assets/tutorial-images/rbison/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width=6,
  fig.height=4,
  cache=TRUE
)
```

`rbison` is an R package to search and retrieve data from the USGS BISON service. `rbison` wraps R code around the BISON API to allow you to talk to the BISON database from R.

BISON has occurrence data for the US only.

BISON is a node of the Global Biodiversity Information Facility (GBIF) - i.e., you can get data that's available in BISON via GBIF instead if you want.

********************

### Info

See [http://bison.usgs.ornl.gov/doc/api.jsp](http://bison.usgs.ornl.gov/doc/api.jsp) for API docs for the BISON API.

********************

<section id="installation">

## Installation

```{r eval=FALSE}
install.packages("rbison")
```

Or development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rbison")
```

```{r}
library("rbison")
```

<section id="usage">

## Usage

Notice that the function `bisonmap` automagically selects the map extent to plot for you, being one of the contiguous lower 48 states, or the lower 48 plus AK and HI, or a global map. If some or all points outside the US, a global map is drawn, and throws a warning. You may want to make sure the occurrence lat/long coordinates are correct.

## Get data

```{r one}
out <- bison(species = "Helianthus annuus", count = 10)
```

Inspect summary

```{r two}
out$summary
```

Map occurrences

```{r}
head(out$counties)
```

## All points within the US (including AK and HI)

Get data

```{r four}
out <- bison(species="Bison bison", count=600)
```

Inspect summary

```{r five}
out$summary
```

Map occurrences

```{r six}
bisonmap(out, tomap="state")
```

## All points within the contiguous 48 states

Get data

```{r seven}
out <- bison(species="Aquila chrysaetos", count=600)
```

Inspect summary

```{r eight}
out$summary
```

Map occurrences

```{r nine}
bisonmap(out, tomap="points")
```

## Plot county or state

With any data returned from a `bison` call, you can choose to plot county or state level data

Counties - using last data call for Aquila

```{r ten}
bisonmap(out, tomap="county")
```

States - using last data call for Aquila

```{r eleven}
bisonmap(out, tomap="state")
```

## Constrain search with county IDs or bounding boxes

### Constrain search to a certain county.

Check out [this site](http://www.epa.gov/enviro/html/codes/state.html) to get state and county fips codes. Fips codes are like so: First two digits are the state code - last three are the county code. For example the *06* in  06037 is the state of California, and the *037* is the Los Angeles county.

```{r countyfips}
out <- bison(species="Helianthus annuus", countyFips = "06037")
```

Inspect summary

```{r}
out$summary
```

By default, the query only returned 10 records

```{r}
head(out$points)
```

Or specify county by its actual name - probably much easier.

```{r countyname}
out <- bison(species="Helianthus annuus", county = "Los Angeles")
```

Inspect summary

```{r}
out$summary
```

By default, the query only returned 10 records

```{r}
head(out$points)
```

`bison` will help you if you spell the name wrong, or use a partial name. The results are not printed below, but you would get a prompt asking you to pick between the two counties that start with *Los*.

```{r countyname_wrong, eval=FALSE}
bison(species="Helianthus annuus", county = "Los")
```

### Constrain search to a amorphous area.

Check out the Wikipedia page [here](http://en.wikipedia.org/wiki/Well-known_text) for an in depth look at the options, terminology, etc.

```{r}
out <- bison(species="Helianthus annuus", aoi = "POLYGON((-111.06360117772908 38.84001566645886,-110.80542246679359 39.37707771107983,-110.20117441992392 39.17722368276862,-110.20666758398464 38.90844075244811,-110.63513438085685 38.67724220095734,-111.06360117772908 38.84001566645886))")
```

Inspect summary

```{r}
out$summary
```

The data

```{r}
out$points
```

### Constrain search to a certain aoibbox.

An aoibbox uses the format minx, miny, maxx, maxy.

```{r boundbox}
out <- bison(species="Helianthus annuus", aoibbox = '-120.31,35.81,-110.57,40.21')
```

Inspect summary

```{r}
out$summary
```

The data, by default, the query only returned 10 records

```{r}
head(out$points)
```


<section id="citing">

## Citing

To cite `rbison` in publications use:

<br>

> Scott Chamberlain (2015). rbison: R interface to the USGS BISON API. R package version 0.4.8 http://CRAN.R-project.org/package=rbison

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rbison](https://github.com/ropensci/rbison/issues?state=open)

[Back to top](#top)
