---
title: rcrossref tutorial
package_version: 0.7.0
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE
)
```



### Installation

```{r eval=FALSE}
install.packages("rcrossref")
```

Or development version from GitHub

```{r eval=FALSE}
devtools::install_github("ropensci/rcrossref")
```

```{r}
library("rcrossref")
```


## Usage

### Citation search

CrossRef's [DOI Content Negotiation](http://www.crosscite.org/cn/) service, where you can citations back in various formats, including `apa`

```{r}
cr_cn(dois = "10.1371/journal.pone.0112608", format = "text", style = "apa")
```

There are a lot more styles. We include a dataset as a character vector within the package, accessible via the `get_styles()` function, e.g.,

```{r}
get_styles()[1:5]
```

`bibtex`

```{r}
cat(cr_cn(dois = "10.1126/science.169.3946.635", format = "bibtex"))
```

`bibentry`

```{r}
cr_cn(dois = "10.6084/m9.figshare.97218", format = "bibentry")
```

### Citation count

Citation count, using OpenURL

```{r}
cr_citation_count(doi="10.1371/journal.pone.0042793")
```

### Search Crossref metadata API

There are two functions that use an older Crossre API: `cr_search()` and `cr_search_free()`. You can of course use them, but the newer Crossref API available through various functions (`cr_agency()`, `cr_fundref()`, `cr_journals()`, `cr_licenses()`, `cr_members()`, `cr_prefixes()`, and `cr_works()`) is more powerful and will recieve more support going forward. The following functions (of the newer set just mentioend) all use the [CrossRef API](https://github.com/CrossRef/rest-api-doc/blob/master/rest_api.md).

#### Funders

```{r}
cr_funders(query="NSF")
```

#### Check the DOI minting agency

```{r}
cr_agency(dois = '10.13039/100000001')
```

#### Search works

```{r}
cr_works(filter=c(has_orcid=TRUE, from_pub_date='2004-04-04'), limit=1)
```

#### Search journals

```{r}
cr_journals(issn=c('1803-2427','2326-4225'))
```

#### Search license information

```{r}
cr_licenses(query = 'elsevier')
```

#### Search based on DOI prefixes

```{r}
cr_prefixes(prefixes=c('10.1016','10.1371','10.1023','10.4176','10.1093'))
```

#### Search CrossRef members

```{r}
cr_members(query='ecology', limit = 5)
```

#### Get N random DOIs

`cr_r()` uses the function `cr_works()` internally.

```{r}
cr_r()
```

You can pass in the number of DOIs you want back (default is 10)

```{r}
cr_r(2)
```

### Citing

To cite `rcrossref` in publications use:

<br>

> Scott Chamberlain, Carl Boettiger, Ted Hart and Karthik Ram (2017).
  rcrossref: Client for Various 'CrossRef' 'APIs'. R package version
  0.7.0. https://CRAN.R-project.org/package=rcrossref


### License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rcrossref](https://github.com/ropensci/rcrossref/issues?state=open)

[Back to top](#top)
