---
title: rdataretriever tutorial
package_version: 1.0.0
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "../assets/tutorial-images/rdataretriever/",
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

<section id="installation">

## Installation

### Install DataRetriever software

To use the R package `rdataretriever` you first need to install `Retriever`. For installation instructions go to <http://www.data-retriever.org/#install>; you can install with `pip`, and there's installers available for all major operating systems, or it can be installed from source.

[Source code](https://github.com/weecology/retriever).

### Install R package

Install and load `rdataretriever` into the R session. Stable version from CRAN

```{r eval=FALSE}
install.packages("rdataretriever")
```

Or development version from Github:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rdataretriever")
```

```{r}
library('rdataretriever')
```

<section id="usage">

## Usage

### List the datasets available via the Retriever

```{r}
rdataretriever::datasets()
```

### Install into csv

Install the Gentry dataset into csv files in your working directory

```{r}
rdataretriever::install(dataset = 'AvianBodySize', connection = 'csv')
head(read.csv("AvianBodySize_species.csv")[,c(1:10)])
```

### Read into R

Install and load a dataset as a list

```{r}
Gentry <- rdataretriever::fetch('Gentry')
```

```{r}
head(Gentry$sites)
```

### Write data into a database

Install a dataset into a SQLite database

```{r}
rdataretriever::install(dataset = 'Gentry', connection = 'sqlite', db_file = "gentrysqlite.sqlite3")
rdataretriever::install(dataset = 'MCDB', connection = 'sqlite', db_file = "mcdbsqlite.sqlite3")
```

Load `RSQLite` and connect to the database

```{r}
library("RSQLite")
db <- dbConnect(SQLite(), "mcdbsqlite.sqlite3")
```

List tables in the database

```{r}
dbListTables(db)
```

List fields in the table

```{r}
dbListFields(db, "MCDB_species")
```

Query and get data

```{r}
dbGetQuery(db, "SELECT * FROM MCDB_species LIMIT 10")
```

<section id="citing">

## Citing

To cite `rdataretriever` in publications use:

Software:

> Daniel McGlinn, Henry Senyondo, Shawn Taylor and Ethan White (2017). rdataretriever: R Interface to the Data Retriever. R package version 1.0.0. https://github.com/ropensci/rdataretriever/

Publication:

> Morris, Benjamin D., and Ethan P. White. (2013). The EcoData Retriever: Improving Access to Existing Ecological Data. PLoS ONE 8 (6) (jun): 65848. doi:10.1371/journal.pone.0065848.

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rdataretriever](https://github.com/ropensci/rdataretriever/issues?state=open)

[Back to top](#top)
