---
title: rdatacite tutorial
package_version: 0.1.0
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "../assets/tutorial-images/rdatacite/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`rdatacite` provides programmatic accesses to [DataCite](http://datacite.org/) metadata

* OAI-PMH. Documentation for the DataCite OAI-PMH is available [here](http://oai.datacite.org/). [Documentation for OAI-PMH](http://www.openarchives.org/OAI/openarchivesprotocol.html)
* REST API. [Docs](http://search.datacite.org/help.html). [Metadata schema](http://schema.datacite.org/). [Solr config file](https://github.com/datacite/search/blob/master/src/main/resources/solrconfig.xml)

<section id="installation">

## Installation

```{r eval=FALSE}
install.packages("rdatacite")
```

Or development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rdatacite")
```

```{r}
library("rdatacite")
```

<section id="usage">

## Usage


### OAI-PMH

#### Ping the service

```{r}
dc_oai_identify()
```

#### List metadata formats

```{r}
dc_oai_listmetadataformats(id = "56225")
```

#### List identifiers

```{r}
dc_oai_listidentifiers(from = '2011-06-01T', until = '2011-07-01T')
```

#### Get records

```{r}
dc_oai_getrecord(id = "56225")
```

You can pass in more than one identifier to `id` parameter.

### Search API

This is the API for the same query interface you'd use on the DataCite website.

#### Search

Search for the term _laser_

```{r}
dc_search(q = "laser", fl = c('doi','publicationYear'), rows = 5)
```

Another search: published between 2000 and 2005

```{r}
dc_search(q = "publicationYear:[2000 TO 2005]", fl = c('doi', 'publicationYear'), rows = 5)
```

#### Facet

```{r}
dc_facet(q = "wind", facet.field = 'publisher_facet', facet.limit = 5)
```

#### Stats


```{r}
dc_stats(q = "ecology", stats.field = 'date')
```

#### More-like-this

```{r}
dc_mlt(q = "ecology", mlt.fl = 'title', mlt.count = 2, fl = 'doi')
```


<section id="citing">

## Citing

To cite `rdatacite` in publications use:

<br>

> Scott Chamberlain (2016). rdatacite: 'DataCite' Client for 'OAI-PMH'
  Methods and their Search 'API'. R package version 0.1.0. https://cran.rstudio.com/package=rdatacite

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rdatacite](https://github.com/ropensci/rdatacite/issues?state=open)

[Back to top](#top)
