---
title: rdpla tutorial
package_version: 0.1.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/rdpla/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`rdpla`: R client for Digital Public Library of America

[Digital Public Library of America][dpla] brings together metadata from libraries, archives,
and museums in the US, and makes it freely available via their web portal as well as
an API. DPLA's portal and API don't provide the items themselves from contributing
institutions, but they provide links to make it easy to find things. The kinds of
things DPLA holds metadata for include images of works held in museums, photographs
from various photographic collections, texts, sounds, and moving images.

DPLA has [a great API](https://github.com/dpla/platform) with good documentation -
a rare thing in this world. Further documentation on their API can be found on their [search fields](https://dp.la/info/developers/codex/responses/field-reference/) and [examples of queries](https://dp.la/info/developers/codex/requests/).  Metadata schema information [here](https://dp.la/info/wp-content/uploads/2013/04/DPLA-MAP-V3.1-2.pdf).

DPLA API has two main services (quoting from [their API docs](https://dp.la/info/developers/codex/requests/)):

* items: A reference to the digital representation of a single piece of content indexed by
the DPLA. The piece of content can be, for example, a book, a photograph, a video, etc. The
content is digitized from its original, physical source and uploaded to an online repository.
items are available in the `dpla_items()` function.
* collections: A collection is a little more abstract than an item. Where an item is a
reference to the digital representation of a physical object, a collection is a
representation of the grouping of a set of items. collections are available in the
`dpla_collections()` function.

Note that you can only run examples/vignette/tests if you have an API key. See below for an example of how
to get an API key.

<section id="installation">

## Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("rdpla")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropensci/rdpla")
```

```{r}
library("rdpla")
```

<section id="usage">

## Usage

### API key

If you already have a DPLA API key, make sure it's in your `.Renviron` or `.Rprofile` file.

If you don't have a DPLA API key, use the `dpla_get_key()` function to get a key.
You only need a valid email address to get a key, for example:

```{r eval=FALSE}
dpla_get_key(email = "foo@bar.com")
#> API key created and sent via email. Be sure to check your Spam folder, too.
```

### Search - items

> Note: limiting fields returned for readme brevity.

Basic search

```{r}
dpla_items(q="fruit", page_size=5, fields=c("provider","creator"))
```

Limit fields returned

```{r}
dpla_items(q="fruit", page_size = 10, fields=c("publisher","format"))
```

Limit records returned

```{r}
dpla_items(q="fruit", page_size=2, fields=c("provider","title"))
```

Search by date

```{r}
dpla_items(q="science", date_before=1900, page_size=10, fields=c("id","date"))
```

Search on specific fields

```{r}
dpla_items(description="obituaries", page_size=2, fields="description")
```

```{r}
dpla_items(subject="yodeling", page_size=2, fields="subject")
```

```{r}
dpla_items(provider="HathiTrust", page_size=2, fields="provider")
```

Spatial search, across all spatial fields

```{r}
dpla_items(sp='Boston', page_size=2, fields=c("id","provider"))
```

Spatial search, by states

```{r}
dpla_items(sp_state='Massachusetts OR Hawaii', page_size=2, fields=c("id","provider"))
```

Faceted search

```{r}
dpla_items(facets=c("sourceResource.spatial.state","sourceResource.spatial.country"),
      page_size=0, facet_size=5)
```

### Search - collections

Search for collections with the words _university of texas_

```{r}
dpla_collections(q="university of texas", page_size=2)
```

You can also search in the `title` and `description` fields

```{r}
dpla_collections(description="east")
```

### Visualize

Visualize metadata from the DPLA - histogram of number of records per state (includes __states__ outside the US)

```{r tidy=FALSE, fig.width=8, fig.height=6}
out <- dpla_items(facets="sourceResource.spatial.state", page_size=0, facet_size=25)
library("ggplot2")
library("scales")
ggplot(out$facets$sourceResource.spatial.state$data, aes(reorder(term, count), count)) +
  geom_bar(stat="identity") +
  coord_flip() +
  theme_grey(base_size = 16) +
  scale_y_continuous(labels = comma) +
  labs(x="State", y="Records")
```


<section id="citing">

## Citing

> Scott Chamberlain (2016). rdpla: Client for the Digital Public Library of America ('DPLA'). R package version 0.1.0. https://github.com/ropensci/rdpla


<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for rdpla](https://github.com/ropensci/rdpla/issues?state=open)


[Back to top](#top)

[dpla]: https://dp.la
