---
title: rebird tutorial
package_version: 0.3.0
---

```{r echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/geojsonio/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
	output.lines = 1:10
)
```

A programmatic interface to the eBird database. Find out more about eBird at [their website](http://ebird.org/content/ebird/).

<section id="installation">

## Installation


You can install the stable version from CRAN

```{r eval=FALSE}
install.packages("rebird")
```

Or the development version from Github

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rebird")
```

Then load the package into the R sesssion

```{r}
library("rebird")
```

<section id="usage">

## Usage

### Sightings at location determined by latitude/longitude

Search for bird occurrences by latitude and longitude point

```{r ebirdgeo1}
ebirdgeo(species = 'spinus tristis', lat = 42, lng = -76)
```

Same, but with additional parameter settings, returning only 10 records, including provisional records, and hotspot records.

```{r ebirdgeo2}
ebirdgeo(lat = 42, lng = -76, max = 10, includeProvisional = TRUE, hotspot = TRUE)
```


### Recent sightings frm location IDs

Search for bird occurrences for two locations by their IDs

```{r ebirdloc1}
ebirdloc(locID = c('L99381','L99382'))
```

Search by location ID and species name, as well as some additional parameter settings

```{r ebirdloc2}
ebirdloc(locID = 'L99381', species = 'larus delawarensis', max = 10, provisional = TRUE, hotspot=TRUE)
```


### Recent observations at a region

Search for bird occurrences by region and species name

```{r ebirdregion1}
ebirdregion(region = 'US', species = 'Setophaga caerulescens')
```

Search by location ID and species name, as well as some additional parameter settings. Note that we use `US-OH` to represent Ohio within the US. [See possible region values](https://confluence.cornell.edu/display/CLOISAPI/eBird-1.1-RegionCodeReference).

```{r ebirdregion2}
ebirdregion(region = 'US-OH', max = 10, provisional = TRUE, hotspot = TRUE)
```


### Recent observations at hotspots

Search for bird occurrences by region and species name

```{r ebirdhotspot}
ebirdhotspot(locID = c('L99381','L99382'), species = 'larus delawarensis')
```


### Frequency of observations at hotspots or regions

Obtain historical frequencies of bird occurrences at a given hotspot

```{r ebirdfreq1}
ebirdfreq(loctype = 'hotspots', loc = 'L196159')
```

Same, but in wide format (for making bar charts)

```{r ebirdfreq2}
ebirdfreq(loctype = 'hotspots', loc = 'L196159', long = FALSE)
```

Obtain frequency data for a given state

```{r ebirdfreq3}
ebirdfreq(loctype = 'states', loc = 'CA-BC')
```

Or county

```{r ebirdfreq4}
ebirdfreq(loctype = 'counties', loc = 'CA-BC-GV')
```

Obtain frequency data within a range of years and months

```{r ebirdfreq5}
ebirdfreq(loctype = 'hotspots', loc = 'L196159', startyear = 2010,
          endyear = 2014, startmonth = 1, endmonth = 3)
```


### Recent notable sightings

Search for notable sightings at a given latitude and longitude

```{r ebirdnotable}
ebirdnotable(lat = 42, lng = -70)
```


### eBird taxonomy

Returns a data.frame of all species in the eBird taxonomy for the given parameter inputs

```{r ebirdtax1}
ebirdtaxonomy()
```

Search for hybrid species only

```{r ebirdtax2}
ebirdtaxonomy(cat="hybrid")
```


### Check eBird region

Check if region is valid in eBird database

```{r ebirdregioncheck}
ebirdregioncheck(loctype = 'counties', loc = 'CA-BC-GV')
```



<section id="citing">

## Citing

To cite `rebird` in publications use:

<br>

> Rafael Maia, Scott Chamberlain, Andy Teucher and Sebastian Pardo (2016). rebird: R Client for the eBird Database of Bird Observations. R package version 0.3.0. https://github.com/ropensci/rebird

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rebird](https://github.com/ropensci/rebird/issues?state=open)

[Back to top](#top)
