---
title: rgeospatialquality tutorial
package_version: 0.2.0
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/rgeospatialquality/",
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`rgeospatialquality` provides R native access to the methods of the [Geospatial Data Quality REST API](https://github.com/VertNet/api-geospatial/blob/master/GeospatialQuality.md). With this API, users can perform some basic quality checks on the geospatial aspect of biodiversity data.

<section id="installation">

## Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("rgeospatialquality")
```

Development version from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropenscilabs/rgeospatialquality")
```

```{r}
library("rgeospatialquality")
```

<section id="usage">

## Usage

### Get quality flags for single records

There are two ways to assess single records and get information on their spatial quality: by providing a list-type object with named elements or by passing the required data via function arguments. In any case, flags are calculated with the function `parse_record`, and the result is a named list with the quality information.

The API works on four specific fields, which should be present to provide the most complete answer: `decimalLatitude`, `decimalLongitude`, `countryCode` and `scientificName`. None of them is mandatory, but the more complete the provided information, the better the result set will be. [See the API documentation](https://github.com/VertNet/api-geospatial/blob/master/GeospatialQuality.md).

#### Passing a record

```{r single_record, eval=FALSE}
rec <- list(decimalLatitude=42.1833,
            decimalLongitude=-1.8332,
            countryCode="ES",
            scientificName="Puma concolor")

parse_record(record=rec)
```

#### Passing individual values as arguments

```{r single_parameters, eval=FALSE}
parse_record(decimalLatitude=42.1833,
             decimalLongitude=-1.8332,
             countryCode="ES",
             scientificName="Puma concolor")
```

#### Structure of response

The response is a list of named elements, each element being the result of a single test. For more info on the meaning of these flags, please [check out the API documentation](https://github.com/VertNet/api-geospatial/blob/master/GeospatialQuality.md).

This is what any of the two calls above would return:

```{r single_flags, echo=FALSE}
parse_record(decimalLatitude=42.1833,
      decimalLongitude=-1.8332,
      countryCode="ES",
      scientificName="Puma concolor")
```


### Get quality flags for sets of more than one record

Apart from assessing records one by one, the API also allows sending a set of records to evaluate them all with a single call, using the `add_flags` function. Records must be provided in the form of a `data.frame`. Just as before, each record should have the four key fields (`decimalLatitude`, `decimalLongitude`, `countryCode` and `scientificName`) to give a response as accurate as possible, although none is mandatory. This time, however, the function returns the provided `data.frame` with a new column, called `flags`, consisting of a list of all geospatial quality assessment results.

```{r multi_dataframe, eval=FALSE}
rec1 <- list(decimalLatitude=42.1833, decimalLongitude=-1.8332, countryCode="ES", scientificName="Puma concolor", ...)
rec2 <- list(...)
...
df <- rbind(rec1, rec2, ...)
df2 <- add_flags(df)
```

One easy way to directly get occurrences with the right format is to use the `occ_data` function in ROpenSci's `rgbif` package ([https://github.com/ropensci/rgbif](https://github.com/ropensci/rgbif)). There is a vignette (`rgbif-synergy`) illustrating how to integrate the two packages to improve the workflow of biodiversity data analysis.


<section id="citing">

## Citing

> Javier Otegui (2016). rgeospatialquality: Wrapper for the Geospatial Data Quality REST API. R package
  version 0.3.2. https://cran.rstudio.com/package=rgeospatialquality


<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for rgeospatialquality](https://github.com/ropenscilabs/rgeospatialquality/issues?state=open)


[Back to top](#top)
