---
title: rinat tutorial
package_version: 0.1.4
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.path="../assets/tutorial-images/rinat/",
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

`rinat` is an R wrapper for iNaturalist APIs for accessing the observations. Detailed documentation of the API is available on [iNaturlaist website](http://www.inaturalist.org/pages/api+reference) and is part of our larger species occurence searching package [SPOCC](http://github.com/ropensci/spocc)

<section id="installation">

## Installation

Stable version from CRAN

```{r eval=FALSE}
install.packages("rinat")
```

Or development version from CRAN

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("rinat", "ropensci")
```

Load the package

```{r}
library("rinat")
```

<section id="usage">

## Usage


### Get observations

__Searching__

_Fuzzy search_

You can search for observations either by common or latin name.  It will search the entire iNaturalist entry, so the search below will return all entries that mention Monarch butterflies, not just entries for Monarchs.

```{r, eval=FALSE}
butterflies <- get_inat_obs(query = "Monarch Butterfly")
head(butterflies[,c(4:6)])
```



```{r message = FALSE}
library("ggplot2")
```


Another use for a fuzzy search is searching for a common name or habitat, e.g. searching for all observations that might occur in a vernal pool.  We can then see all the species names found.
```{r}
vp_obs <- get_inat_obs(query = "vernal pool")
head(vp_obs$species_guess)
```


_Taxon query_
To return only records for a specific species or taxonomic group, use the taxon option.

```{r,eval = FALSE}
## Return just observations in the family Plecoptera
stone_flies <- get_inat_obs(taxon = "Plecoptera")

## Return just Monarch Butterfly records
just_butterflies <- get_inat_obs(taxon = "Danaus plexippus")
```


_Bounding box search_

You can also search within a bounding box by giving a simple set of coordinates.

Search by area

```{r, eval = FALSE}
bounds <- c(38.44047,-125,40.86652, -121.837)
deer <- get_inat_obs(query = "Mule Deer", bounds = bounds)
```

__Other functions__



_Get information and observations by project_

You can get all the observations for a project if you know its ID or name as an intaturalist slug

Just get info about a project

```{r}
vt_crows <- get_inat_obs_project("crows-in-vermont", type = "info", raw = FALSE)
```

Now get all the observations for that project

```{r}
vt_crows_obs <- get_inat_obs_project(vt_crows$id, type="observations")
head(vt_crows_obs[,c(1:4)])
```



_Get observation details_

Detailed information about a specific observation can be retrieved by observation ID.  The easiest way to get the ID is from a more general search.

```{r}
m_obs <- get_inat_obs(query = "Monarch Butterfly")
head(get_inat_obs_id(m_obs$id[1]))
```

_Get all observations by user_

If you just want all the observations by a user you can download them with a user ID.  A word of warning though, this can be quite large (easily into the 1000s)

```{r}
m_obs <- get_inat_obs(query="Monarch Butterfly")
head(get_inat_obs_user(as.character(m_obs$user_login[1]),maxresults=20))[,1:5]
```

_Stats by taxa_

Basic statistics are available for taxa counts by date, date range, place ID (numeric ID), or user ID (string)

By date

```{r}
(counts <- get_inat_taxon_stats(date = "2010-06-14"))
```

By place_ID

```{r}
vt_crows <- get_inat_obs_project("crows-in-vermont", type="info", raw = FALSE)
place_counts <- get_inat_taxon_stats(place = vt_crows$place_id)
place_counts$total
```

_Stats by user_

Similar statistics can be retrived for users.  The same input parameters can be used, but results returned are the top five users by species count and observation count.

By date

```{r}
counts <- get_inat_user_stats(date = "2010-06-14")
counts$most_observations$user[,c(1:3)]
```

By place_ID

```{r}
vt_crows <- get_inat_obs_project("crows-in-vermont", type = "info",raw = FALSE)
place_counts <- get_inat_user_stats(place = vt_crows$place_id)
head(place_counts$most_species$user[,c(1:3)])
```



<section id="citing">

## Citing

To cite `rinat` in publications use:

<br>

> Vijay Barve and Edmund Hart (2014). rinat: Access iNaturalist data through APIs. R package version 0.1.4.

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for rinat](https://github.com/ropensci/rinat/issues?state=open)

[Back to top](#top)
