---
title: ritis vignette
package_version: 0.5.0
---

```{r, eval=TRUE, echo=FALSE}
library(knitr)
knitr::opts_chunk$set(
	fig.path = "../assets/tutorial-images/ritis/",
	warning = FALSE,
	message = FALSE,
	comment = "#>"
)
```

An interface to the Integrated Taxonomic Information System (ITIS)
(<http://www.itis.gov>). Includes functions to work with the ITIS REST
API methods (<http://www.itis.gov/ws_description.html>), as well as the
Solr web service (<http://www.itis.gov/solr_documentation.html>).

<section id="installation">

## Installation

```{r eval=FALSE}
install.packages("ritis")
```

Or development version from GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/ritis")
```

<section id="usage">

## Usage

```{r}
library("ritis")
```

### ITIS Solr interface

There are four methods.

* `itis_search()` - Search
* `itis_group()` - Group
* `itis_highlight()` - Hightlight
* `itis_facet()` - Facet

These four methods use the equivalent functions in the package `solrium`, e.g.,
`ritis::itis_search()` uses `solrium::solr_search()`, etc. The `itis_*()` functions
simply use `...` to allow users to pass on parameters to the wrapped `solrium`
functions. So do read the `solrium` docs.

ITIS Solr API docs: [http://www.itis.gov/solr_documentation.html](http://www.itis.gov/solr_documentation.html)

Some examples:

matches only monomials

```{r}
itis_search(q = "nameWOInd:/[A-Za-z0-9]*[%20]{0,0}*/")
```

matches only binomials

```{r}
itis_search(q = "nameWOInd:/[A-Za-z0-9]*[%20]{1,1}[A-Za-z0-9]*/")
```

### ITIS REST API interface

ITIS REST API docs: [http://www.itis.gov/ws_description.html](http://www.itis.gov/ws_description.html)

The following are some example uses. There are many more methods not shown below

-------

Get accepted names for a TSN

```{r}
accepted_names(tsn = 504239)
```

Get common names for a TSN

```{r}
common_names(tsn = 183833)
```

Full hierarchy for a TSN

```{r}
hierarchy_full(tsn = 37906)
```


<section id="citing">

## Citing

> Scott Chamberlain (2016). ritis: Integrated Taxonomic Information System Client. R
  package version 0.5.0. https://github.com/ropensci/ritis


<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our GitHub repo for ritis](https://github.com/ropensci/ritis/issues?state=open)


[Back to top](#top)
