---
title: solr tutorial
package_version: 0.1.6
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "../assets/tutorial-images/solr/",
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

> This package has been rebooted as `solrium` - check out the tutorial at [https://ropensci.org/tutorials/solrium_tutorial.html](https://ropensci.org/tutorials/solrium_tutorial.html)

<section id="installation">

## Installation


More stable version from CRAN

```{r installcran, eval=FALSE}
install.packages("solr")
```

Development version from Github

```{r installgithub, eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropensci/solr")
```

Load

```{r}
library("solr")
```

<section id="usage">

## Usage

The `solr` package is a general purpose R interface to [Solr](http://lucene.apache.org/solr/)

This package only deals with exracting data from a Solr endpoint, not writing data (pull request or holla if you're interested in writing solr data).

### Solr info

+ [Solr home page](http://lucene.apache.org/solr/)
+ [Highlighting help](http://wiki.apache.org/solr/HighlightingParameters)
+ [Faceting help](http://wiki.apache.org/solr/SimpleFacetParameters)
+ [Installing Solr on Mac using homebrew](http://ramlev.dk/blog/2012/06/02/install-apache-solr-on-your-mac/)
+ [Install and Setup SOLR in OSX, including running Solr](http://risnandar.wordpress.com/2013/09/08/how-to-install-and-setup-apache-lucene-solr-in-osx/)

### Quick start

**Define stuff** Your base url and a key (if needed). This example should work. You do need to pass a key to the Public Library of Science search API, but it apparently doesn't need to be a real one.

```{r}
url <- 'http://api.plos.org/search'
key <- 'key'
```

**Search**

```{r}
solr_search(q='*:*', rows=2, fl='id', base=url, key=key)
```

**Facet**

```{r}
solr_facet(q='*:*', facet.field='journal', facet.query=c('cell','bird'), base=url, key=key)
```

**Highlight**

```{r}
solr_highlight(q='alcohol', hl.fl = 'abstract', rows=2, base = url, key=key)
```

**Stats**

```{r}
out <- solr_stats(q='ecology', stats.field=c('counter_total_all','alm_twitterCount'), stats.facet=c('journal','volume'), base=url, key=key)
```

```{r}
out$data
```

```{r}
out$facet
```

**More like this**

`solr_mlt` is a function to return similar documents to the one

```{r}
out <- solr_mlt(q='title:"ecology" AND body:"cell"', mlt.fl='title', mlt.mindf=1, mlt.mintf=1, fl='counter_total_all', rows=5, base=url, key=key)
out$docs
```

```{r}
out$mlt
```

**Parsing**

`solr_parse` is a general purpose parser function with extension methods `solr_parse.sr_search`, `solr_parse.sr_facet`, and `solr_parse.sr_high`, for parsing `solr_search`, `solr_facet`, and `solr_highlight` function output, respectively. `solr_parse` is used internally within those three functions (`solr_search`, `solr_facet`, `solr_highlight`) to do parsing. You can optionally get back raw `json` or `xml` from `solr_search`, `solr_facet`, and `solr_highlight` setting parameter `raw=TRUE`, and then parsing after the fact with `solr_parse`. All you need to know is `solr_parse` can parse

For example:

```{r}
(out <- solr_highlight(q='alcohol', hl.fl = 'abstract', rows=2, base = url, key=key, raw=TRUE))
```

Then parse

```{r}
solr_parse(out, 'df')
```

**Using specific data sources**

*USGS BISON service*

The occurrences service

```{r}
url <- "http://bisonapi.usgs.ornl.gov/solr/occurrences/select"
solr_search(q='*:*', fl=c('decimalLatitude','decimalLongitude','scientificName'), base=url)
```

The species names service

```{r}
url <- "http://bisonapi.usgs.ornl.gov/solr/scientificName/select"
out <- solr_search(q='*:*', base=url, raw=TRUE)
solr_parse(out, "list")$response$docs[1:3]
```


Sweet!  Love me some `solr`


<section id="citing">

## Citing

To cite `solr` in publications use:

<br>

> Scott Chamberlain (2016). solr: General purpose R interface to Solr. R package version 0.1.6. http://CRAN.R-project.org/package=solr

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for solr](https://github.com/solr/rgauges/issues?state=open)

[Back to top](#top)
