---
title: traits tutorial
package_version: 0.2.0
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "../assets/tutorial-images/rnoaa/",
  fig.width = 6,
  fig.height = 4,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

<section id="installation">

## Installation

Install and load `traits` into the R session. Stable version from CRAN

```{r eval=FALSE}
install.packages("traits")
```

Or development version from Github:

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("ropensci/traits")
```

```{r}
library("traits")
```

<section id="usage">

## Usage

### BetyDB

Get trait data for Willow (_Salix_ spp.)

```{r}
(salix <- betydb_search("Salix Vcmax"))
# equivalent:
# (out <- betydb_search("willow"))
```

Summarise data from the output `data.frame`

```{r}
library("dplyr")
salix %>%
  group_by(scientificname, trait) %>%
  mutate(.mean = as.numeric(mean)) %>%
  summarise(mean = round(mean(.mean, na.rm = TRUE), 2),
            min = round(min(.mean, na.rm = TRUE), 2),
            max = round(max(.mean, na.rm = TRUE), 2),
            n = length(n))
```

### NCBI sequence data

Get sequences by id

```{r}
ncbi_byid(ids = "360040093")
```

Get sequences searching by taxonomic name

```{r}
out <- ncbi_searcher(taxa = "Umbra limi", seqrange = "1:2000")
head(out)
```

### GISD invasive species data

> NOTE: we're moving all functions having to do with where species are native/invasive to a new package [https://github.com/ropenscilabs/originr](https://github.com/ropenscilabs/originr)

```{r}
sp <- c("Carpobrotus edulis", "Rosmarinus officinalis")
g_invasive(sp)
```

Or as simplified output

```{r}
g_invasive(sp, simplify = TRUE)
```

### EOL invasive species data

> NOTE: we're moving all functions having to do with where species are native/invasive to a new package [https://github.com/ropenscilabs/originr](https://github.com/ropenscilabs/originr)

```{r}
eol_invasive_('Brassica oleracea', dataset = 'gisd')
```

Another example, with more species, and from

```{r}
eol_invasive_(c('Lymantria dispar','Cygnus olor',
                'Hydrilla verticillata','Pinus concolor'),
              dataset = 'i3n')
```

### EOL's Traitbank data

Note that EOL's Traitbank does not allow us to search by trait.
The only interface we have is to search by taxon page ID.

Searching for _Potus flavus_, page id `328067`

```{r cache=FALSE}
res <- traitbank(pageid = 328067)
res$graph %>%
  select(dwc.measurementtype,
         dwc.measurementvalue) %>%
  filter(!is.na(dwc.measurementvalue))
```

### Coral

Get the species list and their ids

```{r cache=FALSE}
coral_species()
```

Get data by taxon

```{r cache=FALSE}
coral_taxa(80)
```

Get data by trait

```{r cache=FALSE}
coral_traits(105)
```

### Flora Europaea

```{r}
sp <- c("Lavandula stoechas", "Carpobrotus edulis",
        "Rhododendron ponticum", "Alkanna lutea", "Anchusa arvensis")
sapply(sp, fe_native, simplify = FALSE)
```

### Birdlife International

Habitat data

```{r}
birdlife_habitat(22721692)
```

Threats data

```{r}
birdlife_threats(22721692)
```

### Nativity

> NOTE: we're moving all functions having to do with where species are native/invasive to a new package [https://github.com/ropenscilabs/originr](https://github.com/ropenscilabs/originr)

```{r}
sp <- c("Lavandula stoechas", "Carpobrotus edulis", "Rhododendron ponticum",
      "Alkanna lutea", "Anchusa arvensis")
```

Native in the continental USA?

```{r}
sapply(sp, is_native, where = "Continental US", region = "america")
```

Native on Islas Baleares?

```{r}
sapply(sp, is_native, where = "Islas_Baleares", region = "europe")
```

<section id="citing">

## Citing

To cite `traits` in publications use:

<br>

> Scott Chamberlain, Zachary Foster, Ignasi Bartomeus, David LeBauer, and David Harris (2015). traits: Species Trait Data from Around the Web. R package version 0.2.0. https://github.com/ropensci/traits

<section id="license_bugs">

## License and bugs

* License: [MIT](http://opensource.org/licenses/MIT)
* Report bugs at [our Github repo for traits](https://github.com/ropensci/traits/issues?state=open)

[Back to top](#top)
