
---
slug: "treestartr"
title: Generating reasonable starting trees for complex phylogenetic analyses
package_version: 0.1.0
authors:
  - April Wright
date: 2019-06-04
categories: blog
topicid:
tags:
- Software Peer Review
- R
- community
- software
- packages
- treestartr
- phylogeny
- paleontology
- fossils
- divergence time
twitterImg: img/blog-images/<DATE-SLUG>/name-of-image.png
output: 
  md_document:
    variant: markdown_github
    preserve_yaml: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = 'img/blog-images/2019-06-04-treestartr/')
```

If using this template after knitting you'll need to copy the folder 'img/blog-images/2019-06-04-treestartr/' to ' roweb2/themes/ropensci/static/img/blog-images' but the paths in the resulting Markdown will already be valid.

Introduction including outline of the post.

### First awesome section

#### First awesome subsection of the first awesome section

Below is an image (not generated via R Markdown). I like Hugo[^1].

```{r alwayslabelthechunks, echo = FALSE}
blogdown::shortcode("figure",
                    src = "/img/blog-images/<DATE-SLUG>/name-of-image.png",
                    alt = "write an alternative text describing the information contained in the image",
                    width = "500")
```

Below is another image, a plot, so you might see how to add alternative text in this case.


```{r chunkname, fig.cap='alternative text'}
plot(1:10)
```


### Conclusion

[^1]: Hugo! https://gohugo.io/