---
title: Syntax Highlighting
author:
  - Maëlle Salmon
date: '2020-04-05'
slug: syntax-highlighting
categories: []
tags: []
package_version: 0.1.0
description: A very short summary of your post (~ 100 characters)
twitterImg: blog/2019/06/04/post-template/name-of-image.png
output:
  html_document:
    keep_md: yes
---

This is the R Markdown (.Rmd) template for a blog post or tech note. 
To generate your post with just Markdown (.md), use that template instead.

Throughout this template, including the YAML, 
you should change "post-template" to the slug of your post, 
and "2019-06-04" to your publication date.

Save this file under /content/blog/YYYY-MM-DD-slug/index.Rmd in the local copy of your roweb2 fork.

The chunk below ensures your plots will be inserted in the same folder as your Rmd so you can submit the bundle (.Rmd, .md, images) to roweb2.

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "")

# Knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)

# knitr hook to use Hugo highlighting options
knitr::knit_hooks$set(
  source = function(x, options) {
  hlopts <- options$hlopts
    paste0(
      "```r ",
      if (!is.null(hlopts)) {
      paste0("{",
        glue::glue_collapse(
          glue::glue('{names(hlopts)}={hlopts}'),
          sep = ","
        ), "}"
        )
      },
      "\n", glue::glue_collapse(x, sep = "\n"), "\n```\n"
    )
  }
)
```

One example

```{r hl, hlopts = list(linenos='table',hl_lines='[1,"4-6"]',linenostart=199)}
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
```

Another example

```{r hl2, hlopts = list(hl_lines='[1,"4-6"]')}
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
```


One example

```{r hl3, hlopts = list(linenos='table',linenostart=199)}
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
```

