---
title: Trying to Insert Citations
author:
  - Maëlle Salmon
date: '2020-04-04'
slug: trying-to-insert-citations
categories: []
tags:
  - citation
  - bibtex
package_version: 0.1.0
description: A very short summary of your post (~ 100 characters)
output:
  html_document:
    keep_md: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "", comment = "")

knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)
library("RefManageR")

mybib <- ReadBib("refs.bib", check = FALSE) 

cite <- function(key, bib = mybib) {
  RefManageR::NoCite(bib, key)
  paste0("[^", key, "]")
}
```



Through the helpful feedback from the ROpenSci community, the package
has recently passed software review and a supporting article was
recently published in the Journal of Open Source Software `r cite("Lau2020")`,
in which you can find more details about the package. The package is
hosted through the ROpenSci organization on Github, and the package
can be installed using the [devtools](https://devtools.r-lib.org)
library `r cite("R-devtools")` directly from the repository
(https://github.com/ROpenSci/Rclean).

blabla `r cite("R-CodeDepends")`

more blabla `r cite("Pasquier2017")`

more more blabla `r cite("R-reprex")`

oh right cool `r cite("R-drake")` really `r cite("Carata2014")`

```{r bib, echo = FALSE, results = "asis"}
apa_style <- tools::bibstyle("apa", sortKeys = function(refs) seq_along(refs),
    fmtPrefix = function(paper) paste0("[^", attr(paper,"key"), "]:"),
       .init = TRUE)
RefManageR::PrintBibliography(mybib, .opts = list(bib.style = "apa", sorting = ""))
```