---
title: Experimenting Around the Packages Table
author:
  - Maëlle Salmon
date: '2020-04-30'
slug: table
categories: []
tags: []
package_version: 0.1.0
description: A very short summary of your post (~ 100 characters)
twitterImg: blog/2019/06/04/post-template/name-of-image.png
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "")

# knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)

# knitr hook to use Hugo highlighting options
knitr::knit_hooks$set(
  source = function(x, options) {
  hlopts <- options$hlopts
    paste0(
      "```r ",
      if (!is.null(hlopts)) {
      paste0("{",
        glue::glue_collapse(
          glue::glue('{names(hlopts)}={hlopts}'),
          sep = ","
        ), "}"
        )
      },
      "\n", glue::glue_collapse(x, sep = "\n"), "\n```\n"
    )
  }
)
```


```{r table, echo=FALSE, message=FALSE}

cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}

get_name <- function(list) {
  list$package[[1]]
}

get_description <- function(list) {
  cran_unquote(
    trimws(
    toString(
      list$details$description
      )
  )
  )
}

get_data_source <- function(list) {
  source <- 
      cran_unquote(
          trimws(
            toString(
              list$details$data_source
              )
    )
  )
  
  if (source == "NA"){
    return("")
  }
  
  return(source)
}

packages_json <- jsonlite::read_json("pkgs.json")

packages <- tibble::tibble(
  package = purrr::map_chr(packages_json, get_name),
  data_desc = purrr::map_chr(packages_json, get_description),
  data_source = purrr::map_chr(packages_json, get_data_source)
  )

packages <- packages[packages$package != "genderdata",]
packages <- dplyr::bind_rows(
  packages,
  tibble::tibble(package = "tradestatistics",
                 data_source = "https://tradestatistics.io/")
)

citations <- readr::read_tsv("citations_all.tsv")
citations <- citations[!is.na(citations$doi),]
citations$citation <- gsub("https://doi\\.org.*", "", citations$citation)
citations$citation <- gsub("http://doi\\.org.*", "", citations$citation)
citations$citation <- gsub("doi.*", "", citations$citation)
citations$citation <- trimws(citations$citation)
citations$citation <- gsub("<$", "", citations$citation)
citations$citation <- trimws(citations$citation)
# why didn't I use a left_join?!
get_citations <- function(package, citations) {
  if (!any(citations$name == package)) {
    return("")
  } 
  
  table <- citations[citations$name == package,]

  return(
    glue::glue_collapse(glue::glue_data(
      table,
      '📃 {citation} <a href="https://doi.org/{doi}" target="_blank">https://doi.org/{doi}</a>'), 
      sep = "<br>")
  )
}

packages$citations <- NA

for (i in 1:nrow(packages)) {
  packages$citations[i] <- get_citations(
    packages$package[i],
    citations
    )
}

registry <- jsonlite::read_json("https://ropensci.github.io/roregistry/registry.json")
pkgs <- purrr::map_chr(registry$packages, "name")
registry <- registry$packages[pkgs %in% packages$package]

for (i in 1:length(registry)) {
  registry[[i]]$citations <- packages$citations[packages$package == registry[[i]]$name]
  
    registry[[i]]$data_source <- packages$data_source[packages$package == registry[[i]]$name]
    
  registry[[i]]$details <- cran_unquote(gsub("\\\n","", registry[[i]]$details))
  
  registry[[i]]$description <- cran_unquote(registry[[i]]$description)
}

jsonlite::write_json(
  list(packages = registry),
  path = "registry.json",
  auto_unbox = TRUE,
  pretty = TRUE
  )
```

### Table of `r length(registry)` packages

> The table below of `r length(registry)` packages uses JavaScript libraries including DataTable and jQuery. If you have disabled JavaScript from your browser or read from R-Bloggers or a feed reader rather than from <https://ropensci.org>, the table will be a vanilla html table of a random excerpt of 10 packages.[^solutions]

You can expand rows by clicking on the book emoji :closed_book:.
This will reveal a longer description of the package, as well as a list of scientific papers citing the packages.
Click on the book emoji :book: again to collapse the row.

```{r, echo=FALSE}
library("magrittr")
set.seed(42)
registry %>%
  purrr::map(tibble::as_tibble) %>%
  dplyr::bind_rows() %>%
  dplyr::mutate(book = "") %>%
  dplyr::select(name, description, data_source, maintainer) %>%
  dplyr::mutate(description = htmltools::htmlEscape(description)) %>%
  dplyr::sample_n(10) %>%
  knitr::kable(format = "html", table.attr = "class=\"display\" style=\"width:100%;max-height: 3.5em;\" id=\"packagestable\"", escape = FALSE)
```

<!--html_preserve-->
<script src="https://code.jquery.com/jquery-3.5.0.min.js" integrity="sha256-xNzN2a4ltkB44Mc/Jz3pT4iU1cmeR0FkXs4pru/JxaQ=" crossorigin="anonymous"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/showdown/1.8.7/showdown.min.js" integrity="sha256-CKVcPmoyXVeFTOJvk7+k99gKxTMnKIGs9u8RKFQngVk=" crossorigin="anonymous"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/datatables/1.10.19/js/jquery.dataTables.min.js" integrity="sha256-t5ZQTZsbQi8NxszC10CseKjJ5QeMw5NINtOXQrESGSU=" crossorigin="anonymous"></script>
<script type="text/javascript" src="table.js"></script>
<!-- https://datatables.net/forums/discussion/comment/151719/#Comment_151719 -->
<!--/html_preserve-->

[^solutions]: You can also explore [the JSON feeding the table](registry.json). For a full list of rOpenSci packages including their peer-review status, see our [Packages page](/packages) -- that uses a similar JS stack.