---
title: A roundup of R tools for handling BibTeX
author:
  - Maëlle Salmon
date: '2020-05-07'
slug: rmd-citations
categories: []
tags:
  - citations
  - RMarkdown
description: Some tips around the use of citations/references in R Markdown, in particular with RefManageR
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "")

# knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)

# knitr hook to use Hugo highlighting options
knitr::knit_hooks$set(
  source = function(x, options) {
  hlopts <- options$hlopts
    paste0(
      "```r ",
      if (!is.null(hlopts)) {
      paste0("{",
        glue::glue_collapse(
          glue::glue('{names(hlopts)}={hlopts}'),
          sep = ","
        ), "}"
        )
      },
      "\n", glue::glue_collapse(x, sep = "\n"), "\n```\n"
    )
  }
)
```

Our website is based on Markdown content rendered with Hugo.
Markdown content is in some cases knit from R Markdown, but with [less functionality than if one rendered R Markdown to html as in the blogdown default](https://bookdown.org/yihui/blogdown/output-format.html).
In particular, we cannot use the usual BibTex + CSL + Pandoc-citeproc [dance](https://github.com/rstudio/distill/issues/45) to handle a bibliography.
Thankfully, using the rOpenSci package RefManageR, we can still make our own bibliography building from a BibTeX file without formatting references by hand.
In this post we shall present our custom workflow for inserting citations, as well as more mainstream tools.

### Usual citation workflow in Rmd

To repeat information presented in [Nicholas Tierney's excellent online book "R Markdown for Scientists"](https://rmd4sci.njtierney.com/citing-articles-bibliography-styles.html), to get a bibliography one

* references a bibliography style in the document YAML metadata, with entries such as the one below for R,

```bibtex
  @Manual{my-citation-key-for-r,
    title = {R: A Language and Environment for Statistical Computing},
    author = {{R Core Team}},
    organization = {R Foundation for Statistical Computing},
    address = {Vienna, Austria},
    year = {2020},
    url = {https://www.R-project.org/},
  }
```

* inserts citations using their key, e.g. `@my-citation-key-for-r` for the entry above,

* and optionally uses CSL to control the look of the references list; and a bookdown output format to have the bibliography placed somewhere else than the end.

#### Handy R packages for references: citr and knitcitations

Some R tools out there simplify part of the workflow:

* [citr](https://github.com/crsh/citr) by [Frederik Aust](https://github.com/crsh) provides an RStudio add-in to search for references in a .bib file, inserting the key in the document (mimicking tools you could find in Microsoft Word);

* [knitcitations](https://github.com/cboettig/knitcitations) by [Carl Boettiger](/author/carl-boettiger) allows you to not only use the keys, but also DOIs and URLs to cite a paper. E.g. `citet("10.1098/rspb.2013.1372")` will create a citation to `@Boettiger_2013` after querying the web; and `write.bibtex(file = "references.bib")` will allow you to cache entries in a bibliography file.

Both these packages imports [RefManageR](https://docs.ropensci.org/RefManageR/) by [Mathew W. McLean](http://mwmclean.github.io/), a package that _provides tools for importing and working with bibliographic references_, and that has been [peer-reviewed by rOpenSci](https://github.com/ropensci/software-review/issues/119).

### Citation workflow without Pandoc, with RefManageR

Now, for generating Markdown content like this post source, one cannot use the tools presented earlier.
In our blog guide, we explain that citations are footnotes [whose text is APA formatted](https://blogguide.ropensci.org/technical.html#addcitation) and we explain how to [find that citation text for a package or article](https://blogguide.ropensci.org/technical.html#how-to-find-citation-text-for-a-package-or-article).
In most cases, authors are only citing a few references, so that is fine.
However, sometimes authors would like to use an existing .bib file, that they might have e.g. exported from Zotero.

Thankfully, using RefManageR, we can provide a workflow for adding citations from a bib file, by creating functions similar to knitcitations' functions.

We are using the .bib file below,

```{r, results="asis", echo = FALSE, warning = FALSE}
glue::glue_collapse(
  c("```bibtex",
    readLines("refs.bib"),
    "```"),
  sep = "\n")
```

#### Create a bibliography object

We'll start by creating a `BibEntry` object by calling `RefManageR::ReadBib()`: 

```{r readbib}
mybib <- RefManageR::ReadBib("refs.bib", check = FALSE) 
mybib
class(mybib)
```

#### Create and use a cite function

Then in another chunk, we'll create a function creating the citations in the format we need it to have it appear as a footnote: `[^key]`.

```{r createf}
cite <- function(key, bib = mybib) {
  # create a silent reference so RefManageR
  # knows we've used this entry
  RefManageR::NoCite(bib, key)
  # create the string we need in the Markdown document
  paste0("[^", key, "]")
}
```

From the three entries of the BibTeX file, let us cite two of them, what a nice package`r cite("refmanager")` built on top of R`r cite("my-citation-key-for-r")` (we wrote ` cite("refmanager")` and `cite("my-citation-key-for-r")` as inline code).

#### Print the references list

Finally, for the bibliography to appear, we need to add a call to `RefManageR::PrintBibliography()`, after defining our own bibstyle. 
We need each entry to appear in the Markdown file as

```
[^mykey]: <APA string>
```

To get the first part right, we use the `fmtPrefix` of `tools::bibstyle()`.
To get the second part right, after reading through some [R source code](https://github.com/wch/r-source/blob/28f8367a5514cc0d014cf2aa9e7ce909cd2050af/src/library/tools/R/bibstyle.R), we found how to correctly add DOI and URL when present.
Note that in our case the order of appearance of entries within the references list in the Markdown file is not important at all, since Hugo Markdown handler, Goldmark, will reorder them based on order of appearance in the text.

The chunk below was used at the very end of the R Markdown file producing this post.
It came before the definition of another, not citation, footnote.[^footnote]

````markdown
```{r bib, echo = FALSE, results = "asis"} `r ''`
get_doi <- function(paper) {
  if (is.null(paper$doi)) {
    ""
  } else{
    paste0(" , https://doi.org/", paper$doi)
  }
}
apa_style <- tools::bibstyle("apa", sortKeys = function(refs) seq_along(refs),
    fmtPrefix = function(paper) paste0("[^", attr(paper,"key"), "]:"),
    extraInfo = function(paper) paste0(paper$url, get_doi(paper)),
    .init = TRUE)
RefManageR::PrintBibliography(mybib, .opts = list(bib.style = "apa", sorting = ""))
```
````

We run it below without the `results="asis"` option to show what the output looks like.

```{r bib, echo = FALSE}
get_doi <- function(paper) {
  if (is.null(paper$doi)) {
    ""
  } else{
    paste0(" , https://doi.org/", paper$doi)
  }
}
apa_style <- tools::bibstyle("apa", sortKeys = function(refs) seq_along(refs),
    fmtPrefix = function(paper) paste0("[^", attr(paper,"key"), "]:"),
    extraInfo = function(paper) paste0(paper$url, get_doi(paper)),
    .init = TRUE)
RefManageR::PrintBibliography(mybib, .opts = list(bib.style = "apa", sorting = ""))
```

#### Workflow summary

To summarize our workflow

* Instead of referencing the bibliography file in the Document metadata we create an object referring to it in a chunk by calling `RefManageR::ReadBib()`;

* We use a custom-made `cite()` function using `RefManageR::NoCite()` to signal the use of the reference, and string manipulation to add it in the correct format;

* We make the bibliography appear using `RefManageR::PrintBibliography` and `tools::bibstyle()` in a chunk with the `results="asis"` option, chunk that we put exactly where we want the bibliography to appear.

### Do even more with BibTeX files

The workflow presented earlier was a cool example of using RefManageR to use a BibTeX file.
There are other two packages that are worth knowing about for more .bib gymnastics: bib2df and handlr.

#### bib2df: from BibTeX to tibble

[bib2df](https://docs.ropensci.org/bib2df/) by [Philipp Ottolinger](http://www.ottlngr.de/)[^social] is a package that converts bibliography data from .bib to `tibble` and vice versa. 
Like RefManageR, it has been [peer-reviewed by rOpenSci](https://github.com/ropensci/software-review/issues/124).
In the [words of one the reviewers, Adam Sparks](https://github.com/ropensci/software-review/issues/124#issuecomment-308617830), bib2df is _"something that is simple and does one job and does it well."_.

Let's try it on our .bib file.

```{r bib2df}
df <- bib2df::bib2df("refs.bib")
df
bib2df::df2bib(df)
```

Note that RefManageR has a similar export function.


```{r coerce}
library("magrittr")
df2 <- "refs.bib" %>%
  RefManageR::ReadBib() %>%
  as.data.frame()
df2
```

A difference in the formats is for instance the way authors are parsed.

```{r authors}
df$AUTHOR
df2$author
```

`bib2df` even supports separating name, using [`humaniformat`](https://github.com/ironholds/humaniformat/)

```{r humani}
bib2df::bib2df("refs.bib", separate_names = TRUE)$AUTHOR
```

bib2df helps doing fun or serious analyses of reference databases.

<!--html_preserve-->
{{< tweet 887962776806842369 >}}
<!--/html_preserve-->

#### handlr: from BibTeX to RIS, schema.org, etc.

[handlr](https://docs.ropensci.org/handlr/) by [Scott Chamberlain](/author/scott-chamberlain) is less mature but not less useful.
It's _a tool for converting among citation formats_, inspired by Ruby [bolognese library](https://github.com/datacite/bolognese).
It defines an R6 class so it has all-in-one objects, but you could also use [individual functions reading and writing different formats](https://docs.ropensci.org/handlr/reference/index.html).

```{r handlr}
citation <- handlr::HandlrClient$new(x = "refs.bib")
citation
citation$write(format = "ris")
```

At the moment, there are supported readers for CiteProc, RIS, BibTeX, CodeMeta and supported writers for CiteProc, RIS, BibTeX, schema.org, RDF/XML, CodeMeta.
Quite an arsenal for your bibliography conversion needs!

### Conclusion

In this tech note we have explained how to use a BibTeX file in a Markdown file generated from R Markdown without using pandoc-citeproc.
The star of this post was the [RefManageR](https://docs.ropensci.org/RefManageR/) package, that in turns depends on the [bibtex](https://github.com/romainfrancois/bibtex) package and base R functionality to handle bibliographies.
You can learn more about RefManageR functions [on its documentation website](https://docs.ropensci.org/RefManageR/).
We have also mentioned other R packages offering functionality around BibTeX files: [citr](https://github.com/crsh/citr) and [knitcitations](https://github.com/cboettig/knitcitations) for more easily adding references in standard R Markdown documents, [bib2df](https://docs.ropensci.org/bib2df/) and [handlr](https://docs.ropensci.org/handlr/) for converting BibTeX files to other formats.
If you're interested in learning about yet another workflow involving citations in R Markdown documents, have a look at the Distill web framework wrapped in the R package distill, where [citations handling currently happens via JavaScript](https://github.com/rstudio/distill/issues/45).

```{r bib2, echo = FALSE, results = "asis"}
get_doi <- function(paper) {
  if (is.null(paper$doi)) {
    ""
  } else{
    paste0(" , https://doi.org/", paper$doi)
  }
}
apa_style <- tools::bibstyle("apa", sortKeys = function(refs) seq_along(refs),
    fmtPrefix = function(paper) paste0("[^", attr(paper,"key"), "]:"),
    extraInfo = function(paper) paste0(paper$url, get_doi(paper)),
    .init = TRUE)
RefManageR::PrintBibliography(mybib, .opts = list(bib.style = "apa", sorting = ""))
```

[^footnote]: Hello I am the footnote example.
[^social]: Last year Philipp Ottolinger wrote a nice post ["Being a package maintainer or: The social contract"](https://www.ottlngr.de/post/being-a-package-maintainer/) on his personal website.