---
slug: "parzer"
title: "parzer: Parse Messy Geographic Coordinates"
date: 2020-03-19
author:
  - Scott Chamberlain
description: parzer, a new package for parsing messy geographic coordinates.
tags:
  - geospatial
  - parzer
  - software peer review
  - R
output: 
  html_document:
    keep_md: true
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE
)
```

[parzer][] is a new package for handling messy geographic coordinates. 
The first version is now on CRAN, with binaries coming soon hopefully (see
note about installation below). The package recently [completed rOpenSci
review](https://github.com/ropensci/onboarding/issues/341).

parzer motivation
------------------

The idea for this package started with a tweet from [Noam Ross](https://ropensci.org/author/noam-ross/) 
(https://twitter.com/noamross/status/1070733367522590721) about 15 months ago.

The idea being that sometimes you have geographic coordinates in a messy
format, or in many different formats, etc. You can think of it as being
the package for geographic coordinates that [lubridate][] is for dates.

I started off thinking about wrapping a Javascript library with [Jeroen's][jeroen]
[V8][] R package, but then someone showed me or I found (can't remember) [some C++ code][cll]
from back in 2006 that seemed appropriate. I figured I'd go down the C++
track instead of the Javascript track because I figured I could likely get
better performance out of C++ and have slightly less install headaches for users.


Package installation
------------------

The package is on CRAN so you can use `install.packages`

```{r eval=FALSE}
install.packages("parzer")
```

However, since this package requires compilation you probably want a binary.
Binaries are not available on CRAN yet. You can install a binary like

```{r eval=FALSE}
install.packages("parzer", repos = "https://dev.ropensci.org/")
```

```{r}
library(parzer)
```

Check out the package documentation to get started: <https://docs.ropensci.org/parzer/>

<br><br>

Package basics
------------------

The following is a summary of the functions in the package and what they do:

Parse latitude or longitude separately

- parse_lat
- parse_lon

Parse latitudes and longitudes at the same time

- parse_lon_lat

Parse into separate parts of degrees, minutes, seconds

- parse_parts_lat
- parse_parts_lon

Pull out separately degrees, minutes, seconds, or hemisphere

- pz_degree
- pz_minute
- pz_second
- parse_hemisphere

Add/subtract degrees, minutes, seconds

- pz_d
- pz_m
- pz_s

Some examples:

parse latitudes and longitudes

```{r warning=TRUE}
lats <- c("40.123°", "40.123N74.123W", "191.89", 12, "N45 04.25764")
parse_lat(lats)
longs <- c("45W54.2356", "181", 45, 45.234234, "-45.98739874N")
parse_lon(longs)
```

In the above examples you can see there's a mix of valid coordinate
values as well as invalid values. There's a mix of types supported as well.

Sometimes you may want to parse a geographic coordinate into its component
parts; `parse_parts_lat` and `parse_parts_lon` are what you need:

```{r}
x <- c("191.89", 12, "N45 04.25764")
parse_parts_lon(x)
```

Taking a cue from lubridate, we thought it would be useful to make it easier
to add or subtract numbers for coordinates. Three functions help with this:

```{r collapse=TRUE}
pz_d(31)
pz_d(31) + pz_m(44)
pz_d(31) - pz_m(44)
pz_d(31) + pz_m(44) + pz_s(59)
pz_d(-121) + pz_m(1) + pz_s(33)
```

<br><br>

Use cases
------------------

Check out the [parzer use cases][usecases] vignette on the docs site.
[Get in touch](https://github.com/ropensci/parzer/issues) if you have a use case that
might be good to add to that vignette.


Thanks
------------------

Thanks to the reviewers [Maria Munafó](https://github.com/mvickm)
and [Julien Brun](https://github.com/brunj7) for their time invested
in improving the package.

To Do
------------------

There's more to do. We are thinking about [dropping the Rcpp dependency](https://github.com/ropensci/parzer/issues/17),
support [parsing strings that have both latitude and longitude together](https://github.com/ropensci/parzer/issues/3),
[making error messages better](https://github.com/ropensci/parzer/issues/5), and more.

[parzer]: https://github.com/ropensci/parzer
[usecases]: https://docs.ropensci.org/parzer/articles/use_cases.html
[lubridate]: https://cloud.r-project.org/web/packages/lubridate/
[cll]: https://www.codeproject.com/Articles/15659/Longitude-Latitude-String-Parser-and-Formatter
[jeroen]: https://ropensci.org/author/jeroen-ooms/
[V8]: https://cloud.r-project.org/web/packages/V8/
