---
slug: "citecorp"
title: "citecorp: working with open citations"
date: 2019-09-17
author:
  - Scott Chamberlain
topicid: 1832
tags:
- R
- citecorp
- citations
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
)
```

citecorp is a new (hit CRAN in late August) R package for working with data from the
[OpenCitations Corpus][corpus] (OCC).
[OpenCitations](http://opencitations.net/), run by David Shotton and Silvio Peroni,
houses the OCC, an open repository of scholarly citation data
under the very open [CC0 license][CC0]. The [I4OC (Initiative for Open Citations)][i4oc]
is a collaboration between many parties, with the aim of promoting "unrestricted
availability of scholarly citation data". Citation data is available through Crossref,
and available in R via our packages [rcrossref][], [fulltext][] and [crminer][].
Citation data is also available via the [OCC][corpus]; and this OCC data is now
available in R through the new package citecorp.

__How much citation data does the OCC have?__

Quoting the OpenCitations website (as of today): 

> the OCC has ingested the references from 326,743 citing bibliographic
> resources, and contains information about 13,964,148 citation links to 7,565,367
> cited resources

__Why citation data?__ Citations are the links among scholarly works (articles, books,
etc.), leading to many important uses including finding related articles, calculating
article impact, and even use in academic hiring decisions.

__Why open citation data?__ Until recently most citation data has been locked behind
publisher walls. Unfortunately, many publishers see giving away citation data as
losing potential profits. Through the [I4OC][i4oc], many publishers have made 
their citation metadata public, but some of the largest publishers still have not
done so: Elsevier, American Chemical Society, IEEE. Without all citation data being
open, any work that builds on citation data only has a sub-sample of all citations;
you are drawing conclusions about citations from a rather small subset of all
existing citations. Nonetheless, the currently available open citation data is
an important resource; and can be amended with citation data behind paywalls
for those that have access.

## About citecorp and the OCC

OpenCitations created their own identifiers called Open Citation Identifiers (oci), e.g., 

```
020010009033611182421271436182433010601-02001030701361924302723102137251614233701000005090307
```

You are probably not going to be using oci identifiers, but rather DOIs and/or PMIDs 
(PubMed identifier) and/or PMCIDs (PubMed Central identifier) (see the [PubMed Wikipedia entry][pm]
for more). See `?citecorp::oc_lookup` for methods for cross-walking among identifier types.

OpenCitations has a Sparql endpoint for querying their data; you can find that
at <http://opencitations.net/sparql>; we do interface with the OCC Sparql endpoint 
in citecorp, but we don't provide a user interface directly to it in citecorp.

The OCC is also available as [data dumps](https://opencitations.net/download).

<br>

## Links

citecorp source code: <https://github.com/ropenscilabs/citecorp>

citecorp on CRAN: <https://cloud.r-project.org/web/packages/citecorp/>

<br><br>

## Installation

Install from CRAN

```{r eval=FALSE}
install.packages("citecorp")
```

Development version

```{r eval=FALSE}
remotes::install_github("ropenscilabs/citecorp")
```

Load citecorp

```{r}
library(citecorp)
```

<br>

## Converting among identifiers

Three functions are provided for converting among different identifier types;
each function gives back a data.frame containing the url for the article, PMID,
PMCID and DOI:

* `oc_doi2ids`
* `oc_pmid2ids`
* `oc_pmcid2ids`

```{r}
oc_doi2ids("10.1097/igc.0000000000000609")
oc_pmid2ids("26645990")
oc_pmcid2ids("PMC4679344")
```

Under the hood we interact with their Sparql endpoint to do these queries.

## COCI methods

A series of three more functions are meant for fetching references of,
citations to, or metadata for individual scholarly works.

Here, we look for data for the DOI [10.1108/jd-12-2013-0166][paper]

> Peroni, S., Dutton, A., Gray, T. and Shotton, D. (2015), "Setting our 
> bibliographic references free: towards open citation data", Journal of
> Documentation, Vol. 71 No. 2, pp. 253-277.

_Note: If you don't load `tibble` you get normal data.frame's_

```{r}
library(tibble)
doi <- "10.1108/jd-12-2013-0166"
```

references: the works cited within the paper

```{r}
oc_coci_refs(doi)
```
citations: the works that cite the paper

```{r}
oc_coci_cites(doi)
```

metadata: including the ISSN, volumne, title, authors, etc.

```{r}
oc_coci_meta(doi)
```

<br>

## Use cases

There are many example use cases using the OCC already in the literature. Here 
are a few of those, not necessarily using R:

* [Zhu _et al_. 2019. Nine Million Books and Eleven Million Citations: A Study of Book-Based Scholarly Communication Using OpenCitations. arXiv preprint arXiv:1906.06039](https://arxiv.org/abs/1906.06039)
* [Kaminska 2019. PLOS ONE – a case study of citation analysis of research papers based on the data in an open citation index (The OpenCitations Corpus)][kaminska]
* [Simon _et al_. 2019. BioReader: a text mining tool for performing classification of biomedical literature. BMC Bioinformatics, 19(S13). doi:10.1186/s12859-019-2607-x](https://doi.org/10.1186/s12859-019-2607-x) (OpenCitations data used in Fig. 1)
* [Di Iorio _et al_. 2019. Open data to evaluate academic researchers: an experiment with the Italian Scientific Habilitation. arXiv preprint arXiv:1902.03287](https://arxiv.org/abs/1902.03287)

<br>

## To do

- I'd like to vectorize the functions for converting among IDs (e.g., `oc_doi2ids`)
to make them more user friendly. See <https://github.com/ropenscilabs/citecorp/issues/1>
- Contributors! If you'd like to contribute, head on over to [the repo][citecorp] and
get started

<br>

## Get in touch

Get in touch if you have any citecorp questions in the 
[issue tracker](https://github.com/ropenscilabs/citecorp/issues) or the 
[rOpenSci discussion forum](https://discuss.ropensci.org/).



[citecorp]: https://github.com/ropenscilabs/citecorp
[corpus]: http://opencitations.net/corpus
[CC0]: https://creativecommons.org/publicdomain/zero/1.0/
[i4oc]: https://i4oc.org/
[paper]: https://doi.org/10.1108/jd-12-2013-0166
[kaminska]: https://www.researchgate.net/profile/Anna_Kaminska14/publication/320595746_PLOS_ONE_-_a_case_study_of_citation_analysis_of_research_papers_based_on_the_data_in_an_open_citation_index_The_OpenCitations_Corpus/links/5a6491240f7e9b6b8fd9a0e1/PLOS-ONE-a-case-study-of-citation-analysis-of-research-papers-based-on-the-data-in-an-open-citation-index-The-OpenCitations-Corpus.pdf
[rcrossref]: https://github.com/ropensci/rcrossref
[fulltext]: https://github.com/ropensci/fulltext
[crminer]: https://github.com/ropensci/crminer
[pm]: https://en.wikipedia.org/wiki/PubMed
